---
title: "Preprocessing"
output: html_notebook
---

# Setup

```{r}

rm(list = ls())

library(tidyverse)
library(here)
library(Hmisc)
library(ape)
library(rdist)
library(phytools)
library(treeio)
library(Rtsne)
library(tidytree)

select <- dplyr::select

```

# Define Functions

```{r}

# This is for finding the minimum value in a row in the pairwise matrices while ignoring duplicate birds but still allowing legitimate 0 values, to calculate "perceptual isolation" score of the bird. It's not the most elegant but I tried it so many different, less convoluted ways and it kept messing up in bizarre ways that I couldn't debug. This is fine

MinVal <- function(mat) {
  
  # make a dataframe with names as well as unsuffixed names to store the minimum values
  minvals <- data.frame(Names = rownames(mat)) %>% 
    mutate(pure_names = gsub("\\..*", "", Names))
  
  # remove duplicates from the matrix. This is the bit that's slightly cowboy but the more I think about it, the more I reckon it's far less convoluted than it was in the previous iterations
  processed_mat <- mat %>% 
    rownames_to_column("row_names") %>% 
    mutate(row_names = gsub("\\..*","",row_names)) %>% 
    distinct(row_names, .keep_all=TRUE) %>% 
    column_to_rownames("row_names") %>% 
    t() %>% 
    as.data.frame() %>% 
    rownames_to_column("row_names") %>% 
    mutate(row_names = gsub("\\..*","",row_names)) %>% 
    distinct(row_names, .keep_all=TRUE) %>% 
    column_to_rownames("row_names")
  
  # Diagonals set to NA so they will be ignored
  diag(processed_mat) <- NA
  
  # Find the minimum value of each row
  vals <- apply(processed_mat, 1, function (x) min(x, na.rm = TRUE)) %>% 
    as.data.frame() %>% 
    rename(Values = ".") %>% 
    rownames_to_column("pure_names")
  
  # Pop it back into a dataframe and send it on its way
  minvals <- left_join(minvals, vals, by = "pure_names")
  
  return(minvals)
  
}

# This one calculates similarity based on matches and mismatches in a taxonomy. You just give it a dataframe and a list of the columns to compare. The columns should be Order, Family, Genus and Species if you're just comparing birds. Species should be the binomial species name (including both the genus and the species) because the second part of the species name on its own can match between totally unrelated species (e.g., Dromaius novaehollandiae & Larus novaehollandiae)

TaxSimCalc <- function(data, columns_to_compare) {
  
    # Function to take numbers and punctuation out of names - I had to add this because I have sometimes added numbers to the ends of species to make them unique in the datasets.
  clean <- function(str) {
    # Remove punctuation and numbers
    cleaned_str <- gsub("[[:punct:]0-9]", "", str)
    # Convert to lowercase for case-insensitive matching - probably not necessary but just in case
    tolower(cleaned_str)
  }
  
  data <- data %>% drop_na(columns_to_compare)
  
  # Create an empty matrix
  matrix <- matrix(0, nrow = nrow(data), ncol = nrow(data))
  
  # Calculate matches for taxa
  for (i in 1:(nrow(data) - 1)) {
    for (j in (i + 1):nrow(data)) {

      # Clean and preprocess strings for comparison
      cleaned_i <- clean(data[i, columns_to_compare])
      cleaned_j <- clean(data[j, columns_to_compare])
      
      # Count the number of matching values across columns
      matches <- sum(cleaned_i == cleaned_j)
      
      # Store the matches count in the matrix
      matrix[i, j] <- matches
      matrix[j, i] <- matches
    }
    
    # Make the diagonal terms all perfect matches
    matrix[i, i] <- length(columns_to_compare)
  }
  
  # Make it so that perfect match = 0 and perfect mismatch = how many steps up the taxonomic tree you have to get to before they are in the same taxon. (Which in the case of birds species would always be 4 because they're all in the same class - Aves)
  matrix <- length(columns_to_compare) - matrix
  
  # Convert the matrix to a data frame
  df <- as.data.frame(matrix)
  
  # Put the names back in
  colnames(df) <- data$CanonicalName
  df$Species1 <- data$CanonicalName
  
  # Create pairwise list
  list <- df %>% 
    pivot_longer(cols = -contains("Species1"),
                 names_to = "Species2",
                 values_to = "Dist") %>%  
    mutate(s1 = pmin(Species1, Species2),
           s2 = pmax(Species1, Species2)) %>% 
    select(s1, s2, Dist) %>% 
    distinct() %>% 
    filter(s1 != s2) %>% 
    rename(Species1 = s1,
           Species2 = s2) %>% 
    mutate(ScaleDist = scale(Dist)[,1])
  
  return(list(matrix = df, pairs = list))
}

```

# Import Datasets

```{r}

holman_birds <- read.csv(here("preprocessing/data/holman_data.csv"))
# zoe_birds <- read.csv(here("preprocessing/data/bird_data.csv"))

# Anindilyakwa dataset for testing
anindilyakwa_waddy <- read.csv(here("preprocessing/data/anindilyakwa_waddy.csv"))

tlingit_zoe <- read.csv(here("preprocessing/data/tlingit_data.csv"))

zapotec_zoe <- read.csv(here("preprocessing/data/zapotec_data.csv"))

```

## Make Holman dataset useful

```{r}

holman_birds <- holman_birds %>% 
  rename(Language = culture,
         IndexFG = category_index,
         ScientificName = scientific_name) %>% 
  mutate(Language = str_to_title(Language)) 

# Fixing a spelling mistake

holman_birds$ScientificName[holman_birds$ScientificName == "Myiagra ruficoilis"] <-
  "Myiagra ruficollis"

```

# Check how well the Holman and Zoë Anindilyakwa ones line up

```{r}

anindilyakwa_waddy <- anindilyakwa_waddy %>% 
  mutate(CommonNameOG = ifelse(CommonNameWaddy == "",
                               CommonName,
                               CommonNameWaddy),
         ScientificNameOG = ifelse(ScientificNameWaddy == "", 
                                   ScientificName, 
                                   ScientificNameWaddy))%>% 
  rename(FolkGeneric = LanguageName) %>% 
  group_by(FolkGeneric) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  select(CommonName, 
         ScientificName,
         CommonNameOG,
         ScientificNameOG,
         n,
         FolkGeneric)

anindilyakwa_holman <- holman_birds %>% 
  filter(Language == "Anindilyakwa") %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

in_waddy_not_holman <- setdiff(
  anindilyakwa_waddy %>% filter(FolkGeneric != "") %>% select(ScientificNameOG),                                
  anindilyakwa_holman %>% select(ScientificNameOG))

in_holman_not_waddy <- setdiff(
  anindilyakwa_holman %>% select(ScientificNameOG),
  anindilyakwa_waddy %>% filter(FolkGeneric != "") %>% select(ScientificNameOG))

```

There's a mismatch because of how it's indexed Egretta alba, that's totally fine.
There's another one where two birds that definitely have different names are listed as having the same name, that's annoying and I don't know why.

Also, the emu and the bustard are both missing from Holman. I don't know why and it's very weird because they are two of the big distinctive birds.

But, I have confirmed that the dataset is very close to the same as Waddy's original one, transcribed by me from a book that is currently propping up my right elbow, so I feel confident moving forward with this analysis using all of the languages in the dataset. Yay!

Now to align all of the bird names to the taxonomies that we need. These are:
- 1974 Clements taxonomy (for 1974 traditional similarity)
- 2023 Clements taxonomy (for 2023 traditional similarity)
- BirdLife Version 3 taxonomy (for compatibility with BirdTree for phylogenetic similarity)
- 2021 Clements taxonomy (for compatibility with AVONET)
- Canonical Names (based on 2023 Clements but made unique)

# Adding all taxonomies to datasets 

```{r}

# Full Clements 2023 taxonomy for Family and Order
clements_2023 <- read.csv(here("preprocessing/data/clements_2023.csv")) %>% 
  rename(Order2023 = order, 
         ScientificName2023 = scientific.name,
         CommonName2023 = English.name) %>% 
  mutate(Family2023 = str_trim(str_extract(family,"\\b\\w+")))

# Match genuses with families from 1974

GenusMatch1974 <- read.csv(here("preprocessing/data/genus_match_1974.csv"))

```

Just before I do this, I'm going to just check how well the birdtree taxonomy matches the Birdlife V3, because they are actually a bit different (Birdtree recognises 111 species that aren't in the V3)

```{r}

taxonomy_birdtree <- read.csv(here("preprocessing/data/taxonomy_birdtree.csv"))

taxonomy_V3 <- read.csv(here("preprocessing/data/taxonomy_V3.csv")) %>%
  filter(Recognised == "R")

birdtree_extras <- anti_join(taxonomy_birdtree, 
                             taxonomy_V3, 
                             by = c("ScientificNameBT" = "ScientificNameV3"))

```

Okay now this little dataframe contains the 111 birds that are recognised by birdtree and not V3.

## Anindilyakwa

Specific datasets for Anindilyakwa

```{r}

# Birds that should exist in Anindilyakwa territory
species_NT <- read.csv(here("preprocessing/data/species-list_anindilyakwa.csv"))

# Matching 2022 and 2023 Clements names
eBird_2022_2023_NT <- read.csv(here("preprocessing/data/eBird_2022_2023_NT.csv"))

# Matching 1974 and 2023 Clements names - this one isn't specific to Anindilyakwa
eBird_1974_2023 <- read.csv(here("preprocessing/data/eBird_1974_2023.csv"))
changes_74_23 <- eBird_1974_2023

# Matching V3 and 2023 Clements names - also universal

changes_V3_23 <- read.csv(here("preprocessing/data/changes_V3_Clements23.csv"))

# Matching 2021 and 2023 Clements names - also universal

changes_21_23 <- read.csv(here("preprocessing/data/Changes_Clements21_Clements23.csv"))

# 1974 Orders

orders_74 <- read.csv(here("preprocessing/data/clements_1974_orders.csv")) %>% 
  rename(Family1974 = Family,
         Order1974 = Order) %>% 
  mutate(Order1974 = str_trim(Order1974),
         Family1974 = str_trim(Family1974))

```

First I'll merge the Anindilyakwa data I already have into the Holman dataset 

```{r}

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(anindilyakwa_waddy %>% 
              select(ScientificName, 
                     ScientificNameOG, 
                     CommonName), 
            by = "ScientificNameOG",
            multiple = "all")

# Take away the rows with the double-ups of the reef herons

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter((CommonName != "Pacific Reef-Heron (light morph)" & n != 4) | 
         (CommonName != "Pacific Reef-Heron (dark morph)" & n != 1))

# Join in the missing birds

rows_to_add <- anindilyakwa_waddy %>% 
  filter(FolkGeneric != "") %>%
  anti_join(anindilyakwa_holman, by = "ScientificNameOG") %>% 
  mutate(Language = "Anindilyakwa", IndexFG = NA) %>% 
  select(Language, IndexFG, ScientificNameOG, n, ScientificName, CommonName)

anindilyakwa_holman <- rbind(anindilyakwa_holman, rows_to_add)

anindilyakwa_holman$IndexFG[anindilyakwa_holman$ScientificName == "Haematopus longirostris"] <- 
  NA

max_value <- max(anindilyakwa_holman$IndexFG, na.rm = TRUE)

for (i in seq_along(anindilyakwa_holman$IndexFG)) {
  if (is.na(anindilyakwa_holman$IndexFG[i])) {
    anindilyakwa_holman$IndexFG[i] <- max_value + 1
    max_value <- max_value + 1
  }
}

anindilyakwa_holman <- anindilyakwa_holman %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

```

### Clements 2023 

Clements 2022 names are already in the dataset so I just need to change them to 2023.

```{r}

eBirdNT <- eBird_2022_2023_NT %>% 
  filter(str_count(ScientificName2022, "\\S+") == 2,
         !str_detect(ScientificName2022, "[[:punct:]]"))

anindilyakwa_holman <- anindilyakwa_holman %>% 
   rename(ScientificName2022 = "ScientificName",
          CommonName2022 = CommonName) %>%
  left_join(eBirdNT %>%
              select(ScientificName2022,
                     ScientificName2023,
                     CommonName2023),
            by = "ScientificName2022",
            keep = FALSE) %>% 
  mutate(ScientificName2023 = ifelse(is.na(ScientificName2023),
                                     ScientificName2022,
                                     ScientificName2023),
         CommonName2023 = ifelse(is.na(CommonName2023),
                                 CommonName2022,
                                 CommonName2023))

# Now I'm going to check if anything has punctuation or is just blank, because that means I'll have some googling to do.

check <- anindilyakwa_holman %>% 
  filter(ScientificName2023 == "" | str_detect(ScientificName2023, "[[:punct:]]"))

# Four birds need some attention

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2022=="Ninox boobook"] <- 
  "Ninox boobook"
anindilyakwa_holman$CommonName2023[
  anindilyakwa_holman$CommonName2022=="Southern Boobook"] <- 
  "Southern Boobook"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2022=="Ardea intermedia"] <- 
  "Ardea plumifera"
anindilyakwa_holman$CommonName2023[
  anindilyakwa_holman$CommonName2022=="Intermediate Egret"] <- 
  "Plumed Egret"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2022=="Charadrius mongolus"] <- 
  "Anarhynchus mongolus"
anindilyakwa_holman$CommonName2023[
  anindilyakwa_holman$CommonName2022=="Lesser Sand-Plover"] <- 
  "Siberian Sand-Plover"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2022=="Gelochelidon nilotica"] <- 
  "Gelochelidon macrotarsa"
anindilyakwa_holman$CommonName2023[
  anindilyakwa_holman$CommonName2022=="Gull-billed Tern"] <- 
  "Australian Tern"

# Should be all good now!

# Now I will check if there are any birds that do not appear in the territory where the data was collected, as this may identify birds that had had their definitions or names changed.

notin <- anindilyakwa_holman[!(anindilyakwa_holman$ScientificName2023 %in% species_NT$ScientificName), ]

```

Manually checked all of these birdies:

To keep the same: 
 Gallinago hardwickii - Doesn't seem to live here but I'm leaving it in; it lives close by
 Antigone antigone - Doesn't appear on the island but lives close enough by on the mainland
 Gallinula tenebrosa - Appears nearby on mainland
 Platalea flavipes - Appears nearby on mainland
 Ardea pacifica - appears nearby on the mainland
 Falco subniger - doesn't occur there 
 Smicrornis brevirostris - doesn't appear here but is on mainland close by
 Trichoglossus chlorolepidotus - these don't appear anywhere near here and there doesn't seem to be anythingin the dataset they cound be confused for
 Bathilda ruficauda - doesn't appear here or near here 
 
To change:
 Todiramphus chloris - change to Todiramphus sordidus
 Colluricincla megarhyncha - change to Colluricincla woodwardi (changed my mind on this)
 Megapodius freycinet - change to Megapodius reinwardt (it seems like classifying megapodes has always been bit of a mess but these were probably under the same umbrella for a bit)
 Sphecotheres viridis - change to Sphecotheres vieilloti (used to be classed together)
 Lalage sueurii - change to Lalage tricolor
 Myiagra inquieta - change to Myiagra nana
 Rhipidura rufifrons - change to Rhipidura dryas
 Cacomantis flabelliformis - change to Cacomantis pallidus
 Circus assimilis - change to Circus approximans
 Ixobrychus minutus - change to Ixobrychus dubius



```{r}

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Todiramphus chloris"] <- 
  "Todiramphus sordidus"

# anindilyakwa_holman$ScientificName2023[
#   anindilyakwa_holman$ScientificName2023=="Colluricincla megarhyncha"] <- 
#   "Colluricincla woodwardi"
# This was a bad decision

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Megapodius freycinet"] <- 
  "Megapodius reinwardt"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Sphecotheres viridis"] <- 
  "Sphecotheres vieilloti"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Lalage sueurii"] <- 
  "Lalage tricolor"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Myiagra inquieta"] <- 
  "Myiagra nana"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Rhipidura rufifrons"] <- 
  "Rhipidura dryas"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Cacomantis flabelliformis"] <- 
  "Cacomantis pallidus"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Circus assimilis"] <- 
  "Circus approximans"

anindilyakwa_holman$ScientificName2023[
  anindilyakwa_holman$ScientificName2023=="Ixobrychus minutus"] <- 
  "Ixobrychus dubius"

# Now to import all of the common names from the species list

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(species_NT,
            by = c("ScientificName2023" = "ScientificName"),
            keep = FALSE) %>% 
  mutate(CommonName2023 = 
           ifelse(ScientificNameOG %in% notin$ScientificNameOG &
                    !is.na(CommonName),
                  CommonName,
                  CommonName2023))

# Now all of the 2023 scientific names should be as accurate as they can be, and actually indicating the correct bird. So I can go ahead with finding all of the other names now!

# Wait first I need the Family names from 2023

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     Family2023,
                     Order2023),
            by = "ScientificName2023",
            multiple = "all")

```


### Clements 1974

```{r}

# Add 1974 names for things that have just changed their name without changing species concept

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName2023,
                                       ScientificName1974,
                                       CommonName1974),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

anindilyakwa_holman <- anindilyakwa_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

# Filter out the ones that don't have 1974 data to figure out what to do with them

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName2023) %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName1974,
                                       CommonName1974),
            by = ("CommonName1974")) %>% 
   filter(is.na(ScientificName1974) | 
            !(ScientificName1974 == "Rhodospingus cruentus") & 
            !(ScientificName1974 == "Rhipidura setosa"))

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(anindilyakwa_leftovers)

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificName2023) %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName1974,
                                       CommonName1974),
            by=("ScientificName1974"))

# Change Corvus orru to Corvus cecilae (Australian Crow)
# Change Coracina papuensis to Coracina hypoleuca (White-breasted Cuckoo-Shrike)
# Change Ptilinopus regina to Ptilinopus ewingi (Rose-crowned Pigeon)
# Idea for next time - match the common names first and see if that changes things.
# Change Rhipidura dryas to Rhipidura rufifrons (Arafura fantail)

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Rhipidura dryas"] <- 
  "Rhipidura rufifrons"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Corvus orru"] <- 
  "Corvus cecilae"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Coracina papuensis"] <- 
  "Coracina hypoleuca"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Ptilinopus regina"] <- 
  "Ptilinopus ewingi"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Gelochelidon macrotarsa"] <- 
  "Gelochelidon nilotica"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Chrysococcyx minutillus"] <- 
  "Chalcites minutillus"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Anarhynchus mongolus"] <- 
  "Charadrius mongolus"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Anarhynchus ruficapillus"] <-
  "Charadrius alexandrinus"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Ninox boobook"] <-
  "Ninox novaeseelandiae"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Anas superciliosa"] <-
  "Anas poecilorhyncha"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Chalcophaps longirostris"] <-
  "Chalcophaps indica"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Myiagra nana"] <-
  "Seisura inquieta"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Rhipidura phasiana"] <-
  "Rhipidura fuliginosa"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Smicrornis brevirostris"] <-
  "Smicrornis flavescens"

anindilyakwa_leftovers$ScientificName1974[
  anindilyakwa_leftovers$ScientificName2023 == "Trichoglossus rubritorquis"] <-
  "Trichoglossus haematodus"

# Bind it all together

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(anindilyakwa_leftovers) 

# Change a couple more things

anindilyakwa_holman$ScientificName1974[
  anindilyakwa_holman$ScientificName2023 == "Ramsayornis fasciatus"
] <- "Gliciphila fasciatus"
# This came from the Clements book, it's not the same in avibase

anindilyakwa_holman$ScientificName1974[
  anindilyakwa_holman$ScientificName2023 == "Cincloramphus timoriensis"
] <-  "Megalurus timoriensis"

anindilyakwa_holman$ScientificName1974[
  anindilyakwa_holman$ScientificName2023 == "Edolisoma tenuirostre"
] <-  "Coracina tenuirostris" 

anindilyakwa_holman$ScientificName1974[
  anindilyakwa_holman$ScientificName2023 == "Todiramphus sordidus"
] <- "Halcyon chloris"

anindilyakwa_holman$ScientificName1974[
  anindilyakwa_holman$ScientificName2023 == "Haematopus longirostris"
] <- "Haematopus ostralegus"



anindilyakwa_holman <- anindilyakwa_holman %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% select(Genus1974, 
                                      Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, 
                                 Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

Handbook of Birds of the World and BirdLife International Illustrated Checklist of the Birds of the World Version 3 taxonomy - for compatibility with BirdTree

```{r}

# Add V3 names for things that have just changed their name without changing species concept

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                       ScientificNameV3,
                                       CommonNameV3,
                                     FamilyV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

anindilyakwa_holman <- anindilyakwa_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("CommonNameV3"))

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(anindilyakwa_leftovers)

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by=("ScientificNameV3"))

```

Birds to change:
Himantopus leucocephalus = Himantopus himantopus
Anarhynchus mongolus = Charadrius mongolus

Ixobrychus flavicollis - should be okay
Edolisoma tenuirostre = Coracina tenuirostris
Synoicus chinensis = Coturnix chinensis
Rhipidura dryas = Rhipidura rufifrons
Ardea plumifera = Mesophoyx intermedia
Gelochelidon macrotarsa = Sterna nilotica
Chalcophaps longirostris = Chalcophaps indica
Myiagra nana = Myiagra inquieta 
Trichoglossus rubritorquis = Trichoglossus haematodus
Todiramphus sordidus = Todiramphus chloris

```{r}

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Anarhynchus mongolus"] <- 
  "Charadrius mongolus"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Gelochelidon macrotarsa"] <- 
  "Sterna nilotica"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Ixobrychus dubius"] <- 
  "Ixobrychus minutus"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Chalcophaps longirostris"] <- 
  "Chalcophaps indica"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Myiagra nana"] <- 
  "Myiagra inquieta"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Trichoglossus rubritorquis"] <- "Trichoglossus haematodus"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Todiramphus sordidus"] <- 
  "Todiramphus chloris"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Edolisoma tenuirostre"] <- 
  "Coracina tenuirostris"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Synoicus chinensis"] <- 
  "Coturnix chinensis"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Rhipidura dryas"] <- 
  "Rhipidura rufifrons"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Ardea plumifera"] <- 
  "Mesophoyx intermedia"

anindilyakwa_leftovers$ScientificNameV3[
  anindilyakwa_leftovers$ScientificName2023 == "Gelochelidon macrotarsa"] <- 
  "Sterna nilotica"

# Bind it all back together

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(anindilyakwa_leftovers)

to_change <-anindilyakwa_holman$ScientificName2023 %in% birdtree_extras$ScientificNameBT

look <- anindilyakwa_holman[to_change, ]

anindilyakwa_holman$ScientificNameV3[
  anindilyakwa_holman$ScientificName2023 == "Ninox boobook"] <- 
  "Ninox boobook"

```

### Clements 2021

For compatability with AVONET

```{r}

# Add 2021 names for things that have just changed their name without changing species concept
# Just realised I do not need family names. Should go back and changed that for V3

anindilyakwa_holman <- anindilyakwa_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021,
                                       CommonName2021),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

anindilyakwa_holman <- anindilyakwa_holman %>% 
  mutate(ScientificName2021 = ifelse(
    is.na(ScientificName2021),
    ScientificName2023,
    ScientificName2021),
  )

# Pull out the ones that have changed between 2021 and 2023 and match them based on common name

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificName2021 == "" | is.na(ScientificName2021)) %>% 
  select("Language":"FamilyV3") %>%
  mutate(CommonName2021 = CommonName2023) %>% 
  left_join(changes_21_23 %>% select(ScientificName2021,
                                     CommonName2021),
            by = ("CommonName2021"))

# Bind it all back together

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificName2021 != "" & !is.na(ScientificName2021)) %>% 
  rbind(anindilyakwa_leftovers)

# Pull out the changed ones again and match them on scientific name

anindilyakwa_leftovers <- anindilyakwa_holman %>% 
  filter(ScientificName2021 == "" | is.na(ScientificName2021)) %>% 
  select("Language":"FamilyV3") %>%
  mutate(ScientificName2021 = ScientificName2023) %>% 
  left_join(changes_21_23 %>% select(ScientificName2021,
                                     CommonName2021),
            by=("ScientificName2021"))

anindilyakwa_leftovers$ScientificName2021[
  anindilyakwa_leftovers$ScientificName2023 == "Ardea plumifera"] <- 
  "Ardea intermedia"

anindilyakwa_leftovers$ScientificName2021[
  anindilyakwa_leftovers$ScientificName2023 == "Gelochelidon macrotarsa"] <- 
  "Gelochelidon nilotica"

anindilyakwa_leftovers$ScientificName2021[
  anindilyakwa_leftovers$ScientificName2023 == "Anarhynchus mongolus"] <- 
  "Charadrius mongolus"

anindilyakwa_holman <- anindilyakwa_holman %>% 
  filter(ScientificName2021 != "" & !is.na(ScientificName2021)) %>% 
  rbind(anindilyakwa_leftovers)

anindilyakwa <- anindilyakwa_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b"),
         CommonName = CommonName2023) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Innu

```{r}

montagnais_holman <-  holman_birds %>% 
  filter(Language == "Montagnais")  %>% 
  mutate(combined = paste(ScientificName,IndexFG,sep = "_"))

# Delete double Somateria mollissima (there's a subspecies in the dataset)

# to_delete <- duplicated(montagnais_holman$ScientificName) & montagnais_holman$ScientificName == "Somateria mollissima"

unacceptable_duplicates <- which(duplicated(montagnais_holman$combined))

montagnais_holman <- montagnais_holman[-unacceptable_duplicates,] %>% 
  select(-combined) %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

species_can <-
  read.csv(here("preprocessing/data/species-list_montagnais.csv")) %>% 
  filter(ScientificName2023 != "") %>% 
  distinct()

```

### Clements 2023 (Canonical) Names

```{r}

# I have no idea what taxonomy this list of birds is using, so I'm going to cross-reference it with the 2023 list of birds that should be there

notin <- montagnais_holman[!(montagnais_holman$ScientificName %in% species_can$ScientificName2023), ]

# Okay apart from "Anas" being spelled like "Anus" (lol), it seems like some of the names are just old.
# I will have a look at which ones line up with 2023 names and go from there.

montagnais_holman <- montagnais_holman %>% 
  mutate(ScientificName = str_replace(ScientificName, "^Anus\\b", "Anas")) %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     CommonName2023,
                     Family2023,
                     Order2023),
            by = c("ScientificName" = "ScientificName2023"),
            multiple = "all")

# I'll see if any of the remaining ones are the same as in 1974

montagnais_holman <- montagnais_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                     ScientificName2023,
                                     CommonName2023),
            by = c("ScientificName" = "ScientificName1974"),
            multiple = "all")

# Mash together the 2023 Scientific names

montagnais_holman <- montagnais_holman %>% 
  mutate(ScientificName2023 = ifelse((is.na(ScientificName2023) | 
                                        ScientificName2023 == "") &
                                       !is.na(Family2023), 
                                     ScientificName, 
                                     ScientificName2023)) %>% 
  mutate(CommonName = coalesce(CommonName2023.x, 
                               CommonName2023.y)) %>%
  select(-CommonName2023.x, 
         -CommonName2023.y)

# Time to google some birdies one by one

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Capella gallinago"] <- 
  "Gallinago delicata"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Dendroica petechia"] <- 
  "Setophaga petechia"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Anas carolinensis"] <- 
  "Anas crecca"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Oidemia nigra"] <- 
  "Melanitta americana"  ## Watch this later, in 1974 it was Melanitta nigra

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Falco colombarius"] <- 
  "Falco columbarius"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Erolia melanotos"] <- 
  "Calidris melanotos"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Totanus melanoleucus"] <- 
  "Tringa melanoleuca"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Erolia minutilla"] <- 
  "Calidris minutilla"

montagnais_holman$ScientificName[
  montagnais_holman$ScientificName == "Lopipes lobatus"] <- 
  "Lobipes lobatus"  ## Spelling mistake
montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Lobipes lobatus"] <- 
  "Phalaropus lobatus"

montagnais_holman$ScientificName[
  montagnais_holman$ScientificName == "Picoides tridactylis"] <- 
  "Picoides tridactylus"  ## Another spelling mistake
montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Picoides tridactylus"] <- 
  "Picoides dorsalis"  # Changed it because this bird isn't found in this area

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Iridoprocne bicolor"] <- 
  "Tachycineta bicolor"

montagnais_holman$ScientificName[
  montagnais_holman$ScientificName == "Pinicola enudeator"] <- 
  "Pinicola enucleator" ## Spelling mistake
montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Pinicola enucleator"] <- 
  "Pinicola enucleator"

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Plautus alle"] <- 
  "Alle alle"

# Now that they all have their 2023 names, I'll check again to see if everything should be appearing in that area

notin <- montagnais_holman[!(montagnais_holman$ScientificName2023 %in% species_can$ScientificName2023), ]

# Turns out there's just one and it's sometimes grouped together with something that should be there

montagnais_holman$ScientificName2023[
  montagnais_holman$ScientificName == "Circus cyaneus"] <-
  "Circus hudsonius"

# Now to get the Family and Order data

montagnais_holman <- montagnais_holman %>% 
  select(Language,
         IndexFG,
         n,
         ScientificName,
         ScientificName2023) %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  left_join(clements_2023 %>% select(ScientificName2023,
                                     CommonName2023,
                                     Family2023,
                                     Order2023),
            by = "ScientificName2023",
            multiple = "all") %>% 
  rename(CommonName = CommonName2023)
  
```

### Clements 1974

```{r}

# Add 1974 names for things that have just changed their name without changing species concept

montagnais_holman <- montagnais_holman %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName2023,
                                       ScientificName1974),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

montagnais_holman <- montagnais_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

# Filter out the ones that don't have 1974 data and match them based on their Common Names

montagnais_leftovers <- montagnais_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName) %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("CommonName1974"))

# Bind it all back together again

montagnais_holman <- montagnais_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(montagnais_leftovers %>% select(-CommonName1974))

# Now filter out the ones that still don't have 1974 names and see if the 1974 names might match the original scientific names in the dataset

montagnais_leftovers <- montagnais_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificNameOG) %>% 
  left_join(eBird_1974_2023 %>% select(ScientificName1974,
                                       CommonName1974),
            by=("ScientificName1974"))

# Seems like they do. So bind it all back together again and pull in the Family and Order data from 1974

montagnais_holman <- montagnais_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(montagnais_leftovers %>% select(-CommonName1974)) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% 
              select(Genus1974, Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, 
                                 Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

```{r}
# Add V3 names for things that have just changed their name without changing species concept

montagnais_holman <- montagnais_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                     ScientificNameV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

montagnais_holman <- montagnais_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

# Filter out the ones that don't have a V3 scientific name yet, and match them by common name

montagnais_leftovers <- montagnais_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3),
            by = ("CommonNameV3"))

# Bind it all back together again

montagnais_holman <- montagnais_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(montagnais_leftovers %>% select(-CommonNameV3))

# Filter out leftovers again and match by 2023 name

montagnais_leftovers <- montagnais_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("ScientificNameV3"))

montagnais_leftovers$ScientificNameV3[
  montagnais_leftovers$ScientificName2023 == "Gallinago delicata"] <- 
  "Gallinago gallinago"

# Bind it all back together 

montagnais_holman <- montagnais_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(montagnais_leftovers %>% select(-CommonNameV3, 
                                        -FamilyV3))


```

### Clements 2021

```{r}

# Add 2021 names for things that have just changed their name without changing species concept

montagnais_holman <- montagnais_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. And as it turns out there's only one that isn't NA, and upon googling it I realised it's the same too. Awesome, that just saved me a bunch of work.

montagnais_holman <- montagnais_holman %>% 
  mutate(ScientificName2021 = ScientificName2023)

# Make it all neat and pretty

montagnais <- montagnais_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Rangi

```{r}

rangi_holman <-  holman_birds %>% 
  filter(Language == "Rangi")

# Acceptable duplicates
## Alopochen aegyptiaca
## Colius striatus
## Lagonosticta senegala
## Vidua paradisaea 
## Charadrius hiaticula
## Tringa glareola
## Charadrius tricollaris
## Vidua fischeri
## Emberiza tahapisi

unacceptable_duplicates <- which(duplicated(rangi_holman[,c("ScientificName","IndexFG")]))

rangi_holman <- rangi_holman[-unacceptable_duplicates,] %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

species_TZ <-
  read.csv(here("preprocessing/data/species-list_rangi.csv")) %>% 
  filter(ScientificName2023 != "") %>% 
  distinct()

```

### Clements 2023 (Canonical) Names

```{r}
rangi_holman <- rangi_holman %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     CommonName2023,
                     Family2023,
                     Order2023),
            by = c("ScientificName" = "ScientificName2023"),
            multiple = "all")

# I'll see if any of the remaining ones are the same as in 1974

rangi_holman <- rangi_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                     ScientificName2023,
                                     CommonName2023),
            by = c("ScientificName" = "ScientificName1974"),
            multiple = "all")

# Mash together the 2023 Scientific names

rangi_holman <- rangi_holman %>% 
  mutate(ScientificName2023 = ifelse((is.na(ScientificName2023) | 
                                        ScientificName2023 == "") &
                                       !is.na(Family2023), 
                                     ScientificName, 
                                     ScientificName2023)) %>% 
  mutate(CommonName = coalesce(CommonName2023.x, 
                               CommonName2023.y)) %>%
  select(-CommonName2023.x, 
         -CommonName2023.y)

# There are a lot of birds to google this time.

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Threskiornis aethiopica"] <- 
  "Threskiornis aethiopicus"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Egretta intermedia"] <- 
  "Ardea brachyrhyncha"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Treron calva"] <- 
  "Treron calvus"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Hirundo daurica"] <- 
  "Cecropis daurica"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Delichon urbica"] <- 
  "Delichon urbicum"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Prionops plumata"] <- 
  "Prionops plumatus"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Francolinus afer"] <- 
  "Pternistis afer"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Serinus atrogularis"] <- 
  "Crithagra atrogularis"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Egretta alba"] <- 
  "Ardea alba"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Alopochen aegyptiacus"] <- 
  "Alopochen aegyptiaca"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Tringa hypoleucos"] <- 
  "Actitis hypoleucos"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Cursorius chalcopterus"] <- 
  "Rhinoptilus chalcopterus"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Hirundo griseopyga"] <- 
  "Pseudhirundo griseopyga"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Hirundo fuligula"] <- 
  "Ptyonoprogne fuligula"

rangi_holman$ScientificName[
  rangi_holman$ScientificName == "Mirafra rufacinnamonea"] <- 
  "Mirafra rufocinnamomea" ## Spelling error
rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Mirafra rufocinnamomea"] <- 
  "Mirafra rufocinnamomea"

rangi_holman$ScientificName[
  rangi_holman$ScientificName == "Euplectes axillarius"] <- 
  "Euplectes axillaris" ## Spelling error
rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Euplectes axillaris"] <- 
  "Euplectes axillaris"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Vidua paradisea"] <- 
  "Vidua paradisaea"

rangi_holman$ScientificName[
  rangi_holman$ScientificName == "Nectarinia senegalinsis"] <- 
  "Nectarinia senegalinsis"  # Spelling error

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Nectarinia senegalinsis"] <- 
  "Chalcomitra senegalensis"

# Cross-reference to see if any shouldn't be in the area

notin <- rangi_holman[!(rangi_holman$ScientificName2023 %in% species_TZ$ScientificName2023), ]

# There are a few whose species boundaries have been changed.

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Circaetus gallicus"] <- 
  "Circaetus pectoralis"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Balearica pavonina"] <- 
  "Balearica regulorum"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Porphyrio porphyrio"] <- 
  "Porphyrio madagascariensis"

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Hirundo lucida"] <- 
  "Hirundo angolensis"
## I'm guessing that this one was misidentified by the people who compiled the dataset. It's very similar and is found in the area.

rangi_holman$ScientificName2023[
  rangi_holman$ScientificName == "Anthus novaeseelandiae"] <- 
  "Anthus cinnamomeus"

## I'm keeping Cairina moschata in because it's probably domesticated there, same with Gallus gallus.
## I can't explain Calidris alpina
## Keeping in Apus horus because it lives pretty close by.

# Add in the family and order data

rangi_holman <- rangi_holman %>% 
  select(Language,
         IndexFG,
         n,
         ScientificName,
         ScientificName2023) %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  left_join(clements_2023 %>% select(ScientificName2023,
                                     CommonName2023,
                                     Family2023,
                                     Order2023),
            by = "ScientificName2023",
            multiple = "all") %>% 
  rename(CommonName = CommonName2023)
```

### Clements 1974

```{r}

# Add 1974 names for things that have just changed their name without changing species concept

rangi_holman <- rangi_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName2023,
                                       ScientificName1974),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

rangi_holman <- rangi_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

# Filter out the ones that don't have 1974 data and match them based on their Common Names

rangi_leftovers <- rangi_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("CommonName1974"))

# Bind it all back together again

rangi_holman <- rangi_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(rangi_leftovers %>% select(-CommonName1974))

# Now filter out the ones that still don't have 1974 names and see if the 1974 names might match the original scientific names in the dataset

rangi_leftovers <- rangi_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificNameOG) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by=("ScientificName1974"))

# There are a couple still missing names

rangi_leftovers$ScientificName1974[
  rangi_leftovers$ScientificName2023 == "Vidua paradisaea"] <- 
  "Vidua paradisaea"

rangi_leftovers$ScientificName1974[
  rangi_leftovers$ScientificName2023 == "Ptyonoprogne fuligula"] <- 
  "Ptyonoprogne fuligula"

# Bind it all back together again and pull in the Family and Order data from 1974

rangi_holman <- rangi_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(rangi_leftovers %>% select(-CommonName1974)) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% 
              select(Genus1974, Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

```{r}

# Add V3 names for things that have just changed their name without changing species concept

rangi_holman <- rangi_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                     ScientificNameV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

rangi_holman <- rangi_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

# Filter out the ones that don't have a V3 scientific name yet, and match them by common name

rangi_leftovers <- rangi_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3),
            by = ("CommonNameV3"))

# Bind it all back together again

rangi_holman <- rangi_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(rangi_leftovers %>% select(-CommonNameV3))

# Match by scientific name

rangi_leftovers <- rangi_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("ScientificNameV3"))

rangi_leftovers$ScientificNameV3[
  rangi_leftovers$ScientificName2023 == "Porphyrio madagascariensis"] <- 
  "Porphyrio porphyrio"

rangi_leftovers$ScientificNameV3[
  rangi_leftovers$ScientificName2023 == "Ardea brachyrhyncha"] <- 
  "Mesophoyx intermedia"

rangi_leftovers$ScientificNameV3[
  rangi_leftovers$ScientificName2023 == "Anthus cinnamomeus"] <- 
  "Anthus richardi"

rangi_leftovers$ScientificNameV3[
  rangi_leftovers$ScientificName2023 == "Crithagra atrogularis"] <- 
  "Serinus atrogularis"

# Bind it all back together 

rangi_holman <- rangi_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(rangi_leftovers %>% select(-CommonNameV3, -FamilyV3))

```

### Clements 2021

```{r}

# Add 2021 names for things that have just changed their name without changing species concept

rangi_holman <- rangi_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

rangi_holman <- rangi_holman %>% 
  mutate(ScientificName2021 = ifelse(
    is.na(ScientificName2021),
    ScientificName2023,
    ScientificName2021),
  )

# Pull out the ones that have changed between 2021 and 2023 and match them based on common name

rangi_leftovers <- rangi_holman %>% 
  filter(ScientificName2021 == "" | is.na(ScientificName2021)) %>% 
  select("Language":"ScientificNameV3") %>%
  mutate(CommonName2021 = CommonName) %>% 
  left_join(changes_21_23 %>% select(ScientificName2021,
                                     CommonName2021),
            by = ("CommonName2021"))

rangi_leftovers$ScientificName2021[
  rangi_leftovers$ScientificName2023 == "Ardea brachyrhyncha"] <- 
  "Ardea intermedia"

rangi_leftovers$ScientificName2021[
  rangi_leftovers$ScientificName2023 == "Delichon urbicum"] <- 
  "Delichon urbicum"

rangi_leftovers$ScientificName2021[
  rangi_leftovers$ScientificName2023 == "Bubulcus ibis"] <- 
  "Bubulcus ibis"

rangi_holman <- rangi_holman %>% 
  filter(ScientificName2021 != "" & !is.na(ScientificName2021)) %>% 
  rbind(rangi_leftovers %>% select(-CommonName2021))

```

```{r}

# Make it all pretty now

rangi <- rangi_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Saami

```{r}

saami_holman <- holman_birds %>% 
  filter(Language == "Saami")

# Acceptable Duplicates:
## "Larus argentatus"

unacceptable_duplicates <- which(duplicated(saami_holman[, c("ScientificName", "IndexFG")]))

saami_holman <- saami_holman %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

species_SD <-
  read.csv(here("preprocessing/data/species-list_saami.csv")) %>% 
  filter(ScientificName2023 != "") %>% 
  distinct()

```

### Clements 2023 (Canonical) Names

```{r}

 # First I'm just going to fix the substantial number of spelling mistakes

saami_holman$ScientificName[
  saami_holman$ScientificName == "Perisoerus infaustus"] <- 
  "Perisoreus infaustus"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Turdus iliaeus"] <- 
  "Turdus iliacus"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Tringa hypoleuces"] <- 
  "Tringa hypoleucos"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Sterna paradisea"] <- 
  "Sterna paradisaea"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Phalacrocorax carba"] <- 
  "Phalacrocorax carbo"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Dendrocopus minor"] <- 
  "Dendrocopos minor"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Dendrocopus martius"] <- 
  "Dendrocopos martius"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Accipter gentilis"] <- 
  "Accipiter gentilis"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Accipter nisus"] <- 
  "Accipiter nisus"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Falco columbarias"] <- 
  "Falco columbarius"

saami_holman$ScientificName[
  saami_holman$ScientificName == "Haliaetus albicilla"] <- 
  "Haliaeetus albicilla"

saami_holman <- saami_holman %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     CommonName2023,
                     Family2023,
                     Order2023),
            by = c("ScientificName" = "ScientificName2023"),
            multiple = "all")

# I'll see if any of the remaining ones are the same as in 1974

saami_holman <- saami_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                     ScientificName2023,
                                     CommonName2023),
            by = c("ScientificName" = "ScientificName1974"),
            multiple = "all")

```


```{r}

# Mash together the 2023 Scientific names

saami_holman <- saami_holman %>% 
  mutate(ScientificName2023 = ifelse((is.na(ScientificName2023) | 
                                        ScientificName2023 == "") &
                                       !is.na(Family2023), 
                                     ScientificName, 
                                     ScientificName2023)) %>% 
  mutate(CommonName = coalesce(CommonName2023.x, 
                               CommonName2023.y)) %>%
  select(-CommonName2023.x, 
         -CommonName2023.y)

```

```{r}

# Google some birds

saami_holman$ScientificName2023[
  saami_holman$ScientificName == "Delichon urbica"] <- 
  "Delichon urbicum"

saami_holman$ScientificName2023[
  saami_holman$ScientificName == "Cyanosylvia svecica"] <- 
  "Luscinia svecica"

saami_holman$ScientificName2023[
  saami_holman$ScientificName == "Tringa hypoleucos"] <- 
  "Actitis hypoleucos"

saami_holman$ScientificName2023[
  saami_holman$ScientificName == "Dendrocopus minor"] <- 
  "Dryobates minor"

saami_holman$ScientificName2023[
  saami_holman$ScientificName == "Dendrocopos martius"] <- 
  "Dryocopus martius"

```

```{r}

notin <- saami_holman[!(saami_holman$ScientificName2023 %in% species_SD$ScientificName2023), ]

```

```{r}

# Add in the family and order data

saami_holman <- saami_holman %>% 
  select(Language,
         IndexFG,
         n,
         ScientificName,
         ScientificName2023) %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  left_join(clements_2023 %>% select(ScientificName2023,
                                     CommonName2023,
                                     Family2023,
                                     Order2023),
            by = "ScientificName2023",
            multiple = "all") %>% 
  rename(CommonName = CommonName2023)

```

### Clements 1974

```{r}
# Add 1974 names for things that have just changed their name without changing species concept

saami_holman <- saami_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName2023,
                                       ScientificName1974),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

saami_holman <- saami_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

```


```{r}

# Filter out the ones that don't have 1974 data and match them based on their Common Names

saami_leftovers <- saami_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("CommonName1974"))

```

```{r}

# Bind it all back together again

saami_holman <- saami_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(saami_leftovers %>% select(-CommonName1974))

```


```{r}

# Now filter out the ones that still don't have 1974 names and see if the 1974 names might match the original scientific names in the dataset

saami_leftovers <- saami_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificNameOG) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("ScientificName1974"))

```

```{r}

saami_holman <- saami_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(saami_leftovers %>% select(-CommonName1974)) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% 
              select(Genus1974, Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

```{r}

# Add V3 names for things that have just changed their name without changing species concept

saami_holman <- saami_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                     ScientificNameV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

saami_holman <- saami_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

# Filter out the ones that don't have a V3 scientific name yet, and match them by common name

saami_leftovers <- saami_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3),
            by = ("CommonNameV3"))

# Bind it all back together again

saami_holman <- saami_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(saami_leftovers %>% select(-CommonNameV3))

saami_leftovers <- saami_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("ScientificNameV3"))

# Magically they are all fine. Is it a coincidence that this is the European dataset? Bind it all back together 

saami_holman <- saami_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(saami_leftovers %>% select(-CommonNameV3, -FamilyV3))

```

### Clements 2021

```{r}

# Add 2021 names for things that have just changed their name without changing species concept

saami_holman <- saami_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same. And actually I just checked and the other two are the same as well.

saami_holman <- saami_holman %>% 
  mutate(ScientificName2021 = ifelse(
    (is.na(ScientificName2021) | ScientificName2021 == ""),
    ScientificName2023,
    ScientificName2021),
  )

```

```{r}

# Make it all pretty now

saami <- saami_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Tobelo

I think this data was collected on Halmahera Island

```{r}

tobelo_holman <-  holman_birds %>% 
  filter(Language == "Tobelo")

# Error in dataset: in the dataset it puts Meleagris gallopavo and Pelecanus conspicillatus in a group together, in the original data (Taylor 1990) they are not classed together (p. 120)
tobelo_holman$IndexFG[tobelo_holman$ScientificName == "Pelecanus conspicillatus"] <- max(tobelo_holman$IndexFG) + 1

# Acceptable duplicates:
## Lorius roratus (Eclectus roratus)
## Monarcha alecto (Myiagra alecto) x3
## Symposiachrus trivirgatus (Monarcha trivirgatus)
## Myiagra galeata
## Egretta sacra
## Zosterops atriceps

# Unacceptable duplicates
## Centropus goliath (has a subtype name when it's particularly ugly)
## Artamus leucorhynchus (just has two separate names, both of which are only applicable to the one bird)
## Geoffroyus geoffroyi
## Aplonis metallica
## Aplonis mysolensis
## Tanysiptera galatea
## Ptilinopus superbus - actually just needs to be renamed because of an error
## Dicrurus hottentottus - doesn't show up

change <- duplicated(tobelo_holman$ScientificName) & tobelo_holman$ScientificName == "Ptilinopus superbus"
tobelo_holman$ScientificName[change] <- "Ptilinopus granulifrons"

unacceptable_duplicates <- c(which(
  duplicated(
    tobelo_holman[,c("ScientificName", "IndexFG")])),
  which(tobelo_holman$ScientificName == "Dicrurus hottentottus")[1],
  which(tobelo_holman$ScientificName == "Artamus leucorhynchus")[1],
  which(tobelo_holman$ScientificName == "Lorius roratus")[1])


tobelo_holman <- tobelo_holman[-unacceptable_duplicates,] %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

species_MI <-
  read.csv(here("preprocessing/data/species-list_tobelo.csv")) %>% 
  filter(ScientificName2023 != "") %>% 
  distinct()

```

### Clements 2023 (Canonical) Names

```{r}

# Addressing Spelling Mistakes 

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Columba vittiensis"] <- 
  "Columba vitiensis"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Artamus leucorhynchus"] <- 
  "Artamus leucorynchus"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Reinwardtoena reinwardtii"] <- 
  "Reinwardtoena reinwardti"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Eudynamis scolopacea"] <- 
  "Eudynamys scolopacea"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Ptilinopus hyogaster"] <- 
  "Ptilinopus hyogastrus"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Alisterus amboinensus"] <- 
  "Alisterus amboinensis"

tobelo_holman$ScientificName[
  tobelo_holman$ScientificName == "Semeioptera wallacii"] <- 
  "Semioptera wallacii"

# Figure out which ones are the same as 2023

tobelo_holman <- tobelo_holman %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     CommonName2023,
                     Family2023,
                     Order2023),
            by = c("ScientificName" = "ScientificName2023"),
            multiple = "all")

# I'll see if any of the remaining ones are the same as in 1974

tobelo_holman <- tobelo_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                     ScientificName2023,
                                     CommonName2023),
            by = c("ScientificName" = "ScientificName1974"),
            multiple = "all")

# Mash together the 2023 Scientific names

tobelo_holman <- tobelo_holman %>% 
  mutate(ScientificName2023 = ifelse((is.na(ScientificName2023) | 
                                        ScientificName2023 == "") &
                                       !is.na(Family2023), 
                                     ScientificName, 
                                     ScientificName2023)) %>% 
  mutate(CommonName = coalesce(CommonName2023.x, 
                               CommonName2023.y)) %>%
  select(-CommonName2023.x, 
         -CommonName2023.y)

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Monarcha trivirgatus"] <- 
  "Symposiachrus trivirgatus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Nectarinia jugularis"] <- 
  "Cinnyris clementiae"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Pitta erythrogaster"] <- 
  "Erythropitta rufiventris"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Monarcha trivirgatus"] <- 
  "Symposiachrus trivirgatus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Aceros plicatus"] <- 
  "Rhyticeros plicatus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Locustella fasciolata"] <- 
  "Helopsaltes fasciolatus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Rallus philippensis"] <- 
  "Gallirallus philippensis"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Collocalia vanikorensis"] <- 
  "Aerodramus vanikorensis"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Piezorhynchus alecto"] <- 
  "Myiagra alecto"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Nectarinia sericea"] <- 
  "Leptocoma aspasia"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Lorius roratus"] <- 
  "Eclectus roratus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Monarcha pileatus"] <- 
  "Carterornis pileatus"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Megapodius wallacei"] <- 
  "Eulipoa wallacei"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Habroptila wallacii"] <- 
  "Gallirallus wallacii"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Accipiter griseogularis"] <- 
  "Accipiter hiogaster"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Haliaetus leucogaster"] <- 
  "Icthyophaga leucogaster"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Haliaetus leucogaster"] <- 
  "Icthyophaga leucogaster"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Eudynamys scolopacea"] <- 
  "Eudynamys scolopaceus"

notin <- tobelo_holman[!(tobelo_holman$ScientificName2023 %in% species_MI$ScientificName2023), ]

```

Meleagris gallopavo - it's a wild turkey, it's probably there
Same with Gallus gallus - just a normal chicken

```{r}

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Pachycephala pectoralis"] <- 
  "Pachycephala mentalis"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Amaurornis olivacea"] <- 
  "Amaurornis moluccana"

tobelo_holman$ScientificName2023[
  tobelo_holman$ScientificName == "Amaurornis olivacea"] <- 
  "Amaurornis moluccana"

# Add in the family and order data

tobelo_holman <- tobelo_holman %>% 
  select(Language,
         IndexFG,
         n,
         ScientificName,
         ScientificName2023) %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  left_join(clements_2023 %>% select(ScientificName2023,
                                     CommonName2023,
                                     Family2023,
                                     Order2023),
            by = "ScientificName2023",
            multiple = "all") %>% 
  rename(CommonName = CommonName2023)

```

### Clements 1974

```{r}

# Add 1974 names for things that have just changed their name without changing species concept

tobelo_holman <- tobelo_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName2023,
                                       ScientificName1974),
            by = "ScientificName2023",
            multiple = "all")

```

```{r}

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

tobelo_holman <- tobelo_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

```

```{r}

# Filter out the ones that don't have 1974 data and match them based on their Common Names

tobelo_leftovers <- tobelo_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("CommonName1974"))

```

```{r}

# Bind it all back together again

tobelo_holman <- tobelo_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(tobelo_leftovers %>% select(-CommonName1974))

```

```{r}

# Now filter out the ones that still don't have 1974 names and see if the 1974 names might match the original scientific names in the dataset

tobelo_leftovers <- tobelo_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificNameOG) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("ScientificName1974"))

```

```{r}

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Eurystomus azureus"] <- 
  "Eurystomus orientalis"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Centropus bengalensis"] <-   
  "Centropus toulou"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Otus magicus"] <- 
  "Otus manadensis"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Ducula basilica"] <- 
  "Ducula rufigaster"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Accipiter hiogaster"] <- 
  "Accipiter novaehollandiae"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Accipiter hiogaster"] <- 
  "Accipiter novaehollandiae"

tobelo_leftovers$ScientificName1974[
  tobelo_leftovers$ScientificName2023 == "Pitta maxima"] <- 
  "Pitta sordida"

```

Good:
Aerodramus vanikorensis
Myiagra galeata
Eclectus roratus

Dunno:
Pitta maxima
Lonchura molucca

```{r}

tobelo_holman <- tobelo_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(tobelo_leftovers %>% select(-CommonName1974)) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% 
              select(Genus1974, Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

```{r}

# Add V3 names for things that have just changed their name without changing species concept

tobelo_holman <- tobelo_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                     ScientificNameV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

tobelo_holman <- tobelo_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

# Filter out the ones that don't have a V3 scientific name yet, and match them by common name

tobelo_leftovers <- tobelo_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3),
            by = ("CommonNameV3"))

# Bind it all back together again

tobelo_holman <- tobelo_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(tobelo_leftovers %>% select(-CommonNameV3))

tobelo_leftovers <- tobelo_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("ScientificNameV3"))

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Cinnyris clementiae"] <- 
  "Nectarinia jugularis"

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Erythropitta rufiventris"] <- 
  "Pitta erythrogaster"

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Hypsipetes affinis"] <- 
  "Alophoixus affinis"

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Accipiter hiogaster"] <- 
  "Accipiter novaehollandiae"

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Pachycephala mentalis"] <- 
  "Pachycephala pectoralis"

tobelo_leftovers$ScientificNameV3[
  tobelo_leftovers$ScientificName2023 == "Helopsaltes fasciolatus"] <- 
  "Locustella fasciolata"

# Bind it all back together 

tobelo_holman <- tobelo_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(tobelo_leftovers %>% select(-CommonNameV3, -FamilyV3))

```

### Clements 2021

```{r}

# Add 2021 names for things that have just changed their name without changing species concept

tobelo_holman <- tobelo_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same. And besides this one guy who needs changing, all of the blank ones can change too. 

tobelo_holman$ScientificName2021[
  tobelo_holman$ScientificName2023 == "Cinnyris clementiae"] <- 
    "Cinnyris jugularis"


tobelo_holman <- tobelo_holman %>% 
  mutate(ScientificName2021 = ifelse(
    (is.na(ScientificName2021) | ScientificName2021 == ""),
    ScientificName2023,
    ScientificName2021),
  )

```

```{r}

# Make it all pretty now

tobelo <- tobelo_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Tzeltal

```{r}

tzeltal_holman <-  holman_birds %>% 
  filter(Language == "Tzeltal") %>% 
  distinct() %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()

# Acceptable duplicates:
## Aratinga canicularis

unacceptable_duplicates <- which(duplicated(tzeltal_holman[,c("ScientificName", "IndexFG")]))

species_CS <-
  read.csv(here("preprocessing/data/species-list_tzeltal.csv")) %>% 
  filter(ScientificName2023 != "") %>% 
  distinct()

```

### Clements 2023 (Canonical) Names

```{r}

tzeltal_holman <- tzeltal_holman %>% 
  left_join(clements_2023 %>% 
              select(ScientificName2023,
                     CommonName2023,
                     Family2023,
                     Order2023),
            by = c("ScientificName" = "ScientificName2023"),
            multiple = "all")

tzeltal_holman <- tzeltal_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                     ScientificName2023,
                                     CommonName2023),
            by = c("ScientificName" = "ScientificName1974"),
            multiple = "all")

tzeltal_holman <- tzeltal_holman %>% 
  mutate(ScientificName2023 = ifelse((is.na(ScientificName2023) | 
                                        ScientificName2023 == "") &
                                       !is.na(Family2023), 
                                     ScientificName, 
                                     ScientificName2023)) %>% 
  mutate(CommonName = coalesce(CommonName2023.x, 
                               CommonName2023.y)) %>%
  select(-CommonName2023.x, 
         -CommonName2023.y)

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Accipiter chionogaster"] <- 
  "Accipiter striatus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Columba fasciata"] <- 
  "Patagioenas fasciata"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Otus guatemalae"] <- 
  "Megascops guatemalae"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Caprimulgus vociferus"] <- 
  "Antrostomus vociferus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Aratinga holochlora"] <- 
  "Psittacara holochlorus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Centurus aurifrons"] <- 
  "Melanerpes aurifrons"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Thryothorus modestus"] <- 
  "Cantorchilus modestus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Troglodytes musculus"] <- 
  "Troglodytes aedon"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Dendroica petechia"] <- 
  "Setophaga petechia"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Melozone biarcuatum"] <- 
  "Melozone biarcuata"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Chlorospingus ophthalmicus"] <- 
  "Chlorospingus flavopectus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Icterus sclateri"] <- 
  "Icterus pustulatus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Egretta alba"] <- 
  "Ardea alba"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Aulacorynchus prasinus"] <- 
  "Aulacorhynchus prasinus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Piculus rubiginosus"] <- 
  "Colaptes rubiginosus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Myiozetes similis"] <- 
  "Myiozetetes similis"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Vermivora superciliosa"] <- 
  "Oreothlypis superciliosa"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Myioborus picta"] <- 
  "Myioborus pictus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Zarhynchus wagleri"] <- 
  "Psarocolius wagleri"

```

```{r}

notin <- tzeltal_holman[!(tzeltal_holman$ScientificName2023 %in% species_CS$ScientificName2023), ]

```

- Can't really figure out why the Anas platyrhynchos is in there. There are a couple of Anas that aren't but none of them have ever been conspecific with this guy.
- Meleagris gallopavo is a wild turkey, they're just kind of everywhere
- Gallus gallus is in there three times, I'll have to check that.
- Turdus migratorius doesn't appear here, there's an endemic bird that's pretty similar and I think maybe the researchers got them confused at the time

Birds that have been reclassified: 
- Stelgidopteryx ruficollis
- Amazona ochrocephala
- Trogon violaceus
- Momotus momota
- Certhia familiaris
- Contopus cinereus
- Pipilo erythrophthalmus
- Saltator coerulescens
 
```{r}

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Stelgidopteryx ruficollis"] <- 
  "Stelgidopteryx serripennis"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Amazona ochrocephala"] <- 
  "Amazona auropalliata"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Trogon violaceus"] <- 
  "Trogon caligatus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Momotus momota"] <- 
  "Momotus coeruliceps"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Certhia familiaris"] <- 
  "Certhia americana"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Contopus cinereus"] <- 
  "Contopus bogotensis"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Turdus migratorius"] <- 
  "Turdus rufopalliatus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Pipilo erythrophthalmus"] <- 
  "Pipilo maculatus"

tzeltal_holman$ScientificName2023[
  tzeltal_holman$ScientificName == "Saltator coerulescens"] <- 
  "Saltator grandis"

```

```{r}

# Add in the family and order data

tzeltal_holman <- tzeltal_holman %>% 
  select(Language,
         IndexFG,
         n,
         ScientificName,
         ScientificName2023) %>% 
  rename(ScientificNameOG = ScientificName) %>% 
  left_join(clements_2023 %>% select(ScientificName2023,
                                     CommonName2023,
                                     Family2023,
                                     Order2023),
            by = "ScientificName2023",
            multiple = "all") %>% 
  rename(CommonName = CommonName2023)

```

### Clements 1974

```{r}
# Add 1974 names for things that have just changed their name without changing species concept

tzeltal_holman <- tzeltal_holman %>% 
  left_join(changes_74_23 %>% select(ScientificName2023,
                                       ScientificName1974),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between 1974 and 2023. So we can just make them the same

tzeltal_holman <- tzeltal_holman %>% 
  mutate(ScientificName1974 = ifelse(
    is.na(ScientificName1974),
    ScientificName2023,
    ScientificName1974),
  )

# Filter out the ones that don't have 1974 data and match them based on their Common Names

tzeltal_leftovers <- tzeltal_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(CommonName1974 = CommonName) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("CommonName1974"))

# Bind it all back together again

tzeltal_holman <- tzeltal_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(tzeltal_leftovers %>% select(-CommonName1974))

# Now filter out the ones that still don't have 1974 names and see if the 1974 names might match the original scientific names in the dataset

tzeltal_leftovers <- tzeltal_holman %>% 
  filter(ScientificName1974 == "" | is.na(ScientificName1974)) %>% 
  select("Language":"Order2023") %>%
  mutate(ScientificName1974 = ScientificNameOG) %>% 
  left_join(changes_74_23 %>% select(ScientificName1974,
                                             CommonName1974),
            by = ("ScientificName1974"))

tzeltal_leftovers$ScientificName1974[
  tzeltal_leftovers$ScientificName2023 == "Strix fulvescens"] <- 
  "Strix varia"

tzeltal_leftovers$ScientificName1974[
  tzeltal_leftovers$ScientificName2023 == "Aulacorhynchus prasinus"] <- 
  "Aulacorhynchus prasinus"

tzeltal_leftovers$ScientificName1974[
  tzeltal_leftovers$ScientificName2023 == "Icterus galbula"] <- 
  "Icterus bullockii"

tzeltal_holman <- tzeltal_holman %>% 
  filter(ScientificName1974 != "" & !is.na(ScientificName1974)) %>% 
  rbind(tzeltal_leftovers %>% select(-CommonName1974)) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  left_join(GenusMatch1974 %>% 
              select(Genus1974, Family1974), 
            by = "Genus1974") %>% 
  left_join(orders_74 %>% select(Family1974, Order1974), 
            by = "Family1974",
            multiple = "all")

```

### V3

```{r}

# Add V3 names for things that have just changed their name without changing species concept

tzeltal_holman <- tzeltal_holman %>% 
  left_join(changes_V3_23 %>% select(ScientificName2023,
                                     ScientificNameV3),
            by = "ScientificName2023",
            multiple = "all")

# Now the ones that have NA in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same

tzeltal_holman <- tzeltal_holman %>% 
  mutate(ScientificNameV3 = ifelse(
    is.na(ScientificNameV3),
    ScientificName2023,
    ScientificNameV3),
  )

# Filter out the ones that don't have a V3 scientific name yet, and match them by common name

tzeltal_leftovers <- tzeltal_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(CommonNameV3 = CommonName) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3),
            by = ("CommonNameV3"))

# Bind it all back together again

tzeltal_holman <- tzeltal_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(tzeltal_leftovers %>% select(-CommonNameV3))

tzeltal_leftovers <- tzeltal_holman %>% 
  filter(ScientificNameV3 == "" | is.na(ScientificNameV3)) %>% 
  select("Language":"Order1974") %>%
  mutate(ScientificNameV3 = ScientificName2023) %>% 
  left_join(changes_V3_23 %>% select(ScientificNameV3,
                                     CommonNameV3,
                                     FamilyV3),
            by = ("ScientificNameV3"))

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Antrostomus vociferus"] <- 
  "Caprimulgus vociferus"

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Momotus coeruliceps"] <- 
  "Momotus momota"

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Contopus bogotensis"] <- 
  "Contopus cinereus"

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Cantorchilus modestus"] <- 
  "Thryothorus modestus"

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Chlorospingus flavopectus"] <- 
  "Chlorospingus ophthalmicus"

tzeltal_leftovers$ScientificNameV3[
  tzeltal_leftovers$ScientificName2023 == "Saltator grandis"] <- 
  "Saltator coerulescens"

# Bind it all back together 

tzeltal_holman <- tzeltal_holman %>% 
  filter(ScientificNameV3 != "" & !is.na(ScientificNameV3)) %>% 
  rbind(tzeltal_leftovers %>% select(-CommonNameV3, -FamilyV3))
```

### Clements 2021

```{r}
# Add 2021 names for things that have just changed their name without changing species concept

tzeltal_holman <- tzeltal_holman %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Bubo virginianus is fine
# Sturnella magna is fine
# Chlorospingus flavopectus is fine

tzeltal_holman$ScientificName2021[
  tzeltal_holman$ScientificName2023 == "Contopus bogotensis"] <-
  "Contopus cinereus"

# Now the ones that have NA or are blank in the dataset are ones that have not changed between V3 and 2023. So we can just make them the same. 

tzeltal_holman <- tzeltal_holman %>% 
  mutate(ScientificName2021 = ifelse(
    (is.na(ScientificName2021) | ScientificName2021 == ""),
    ScientificName2023,
    ScientificName2021),
  )
```

```{r}

# Make it all pretty now

tzeltal <- tzeltal_holman %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Tlingit

```{r}

tlingit_zoe <- tlingit_zoe %>% 
  mutate(Language = "Tlingit") %>%
  rename(CommonName = CommonName2023) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  filter(!is.na(FolkGeneric)) %>%
  filter(!is.na(ScientificName2023)) %>% 
  distinct(ScientificName2023, .keep_all = TRUE) %>% 
  mutate(FolkGeneric = gsub("\\d+","",FolkGeneric)) %>% 
  mutate(IndexFG = as.integer(factor(FolkGeneric, 
                                     levels = unique(FolkGeneric)))) %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()
  
species_AK <- read.csv(here("preprocessing/data/species-list_tlingit.csv")) %>% distinct(ScientificName2023, .keep_all=TRUE)

```

```{r}

diff <- setdiff(tlingit_zoe$ScientificName2023,species_AK$ScientificName2023)

# Turkey is fine, they're everywhere

```

This dataset already has the current 2023 names, the 1974 taxonomic information and the V3 names because I preprocessed it for something else. Just needs Clements 2021. And also one V3 name that I missed the first time I preprocessed it.

### Clements 2021

```{r}

# Add 2021 names for things that have just changed their name without changing species concept

tlingit_zoe <- tlingit_zoe %>% 
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

tlingit_zoe$ScientificName2021[
  tlingit_zoe$ScientificName2023 == "Accipiter atricapillus"] <- 
  "Accipiter gentilis"

# There's an owl too but it's fine, I checked.

tlingit_zoe <- tlingit_zoe %>% 
  mutate(ScientificName2021 = ifelse(is.na(ScientificName2021) | ScientificName2021 == "",
                                     ScientificName2023, 
                                     ScientificName2021))

```

```{r}

tlingit_zoe$ScientificNameV3[tlingit_zoe$ScientificName2023 == "Pica hudsonia"] <- "Pica hudsonia"

```

```{r}

# Make it all pretty now

tlingit <- tlingit_zoe %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

## Zapotec

```{r}

zapotec_zoe <- zapotec_zoe %>% 
  mutate(Language = "Zapotec") %>%
  rename(CommonName = CommonName2023) %>% 
  mutate(Genus1974 = str_extract(ScientificName1974, "\\b\\w+\\b")) %>% 
  filter(!is.na(FolkGeneric)) %>%
  filter(!is.na(ScientificName2023)) %>% 
  distinct(ScientificName2023, .keep_all = TRUE) %>% 
  mutate(FolkGeneric = gsub("\\d+","",FolkGeneric)) %>% 
  mutate(IndexFG = as.integer(factor(FolkGeneric, 
                                     levels = unique(FolkGeneric)))) %>% 
  group_by(IndexFG) %>% 
  mutate(n = n()) %>% 
  ungroup()
  
species_OA <- read.csv(here("preprocessing/data/species-list_zapotec.csv")) %>% distinct(ScientificName2023, .keep_all=TRUE)

```

```{r}
diff <- setdiff(zapotec_zoe$ScientificName2023,species_OA$ScientificName2023)
print(diff)

# Anser anser seems to be there sometimes maybe?
# Anas platyrhynchos turns up there sometimes
# Anser cygnoides same
# Meleagris gallopavo is that same turkey that's just turning up everywhere
# Gallus gallus is the chicken that is also everywhere.

zapotec_zoe$ScientificName2023[zapotec_zoe$ScientificName2023 == "Ciccaba virgata"] <- 
  "Strix virgata"

```

Like the Tlingit dataset, I have already preprocessed this one a bunch for a different project, so I'll skip V3 and Clements 1974 and just add in Clements 2021

### Clements 2021

```{r}
# Add 2021 names for things that have just changed their name without changing species concept

zapotec_zoe <- zapotec_zoe %>%
  left_join(changes_21_23 %>% select(ScientificName2023,
                                       ScientificName2021),
            by = "ScientificName2023",
            multiple = "all")

# Manually checked and everything is fine

zapotec_zoe <- zapotec_zoe %>%
  mutate(ScientificName2021 = ifelse(is.na(ScientificName2021) | ScientificName2021 == "",
                                     ScientificName2023,
                                     ScientificName2021))

```

### Make it all pretty now

```{r}
zapotec <- zapotec_zoe %>% 
  mutate(Genus2023 = str_extract(
    ScientificName2023, 
    "\\b\\w+\\b")) %>% 
  select(Language,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

```

# Combine all languages into a master dataset

```{r}

bird_data <- rbind(anindilyakwa, 
                   montagnais, 
                   rangi, 
                   saami, 
                   tlingit, 
                   tobelo, 
                   tzeltal, 
                   zapotec) %>% 
  mutate(CanonicalName = make.unique(ScientificName2023, sep = "."),
         IndexFG = paste0(str_sub(Language,1,3), ".", IndexFG)) %>% 
  select(Language,
         CanonicalName,
         CommonName,
         IndexFG,
         n,
         ScientificName2023,
         Genus2023,
         Family2023,
         Order2023,
         ScientificName1974,
         Genus1974,
         Family1974,
         Order1974,
         ScientificName2021,
         ScientificNameV3)

bird_data$CanonicalName <- make.unique(bird_data$ScientificName2023)

# Just check about these extra V3 names

check <- bird_data$ScientificName2023 %in% birdtree_extras$ScientificNameBT
check2 <- bird_data$CommonName %in% birdtree_extras$CommonNameBT

look <- bird_data[check,]
look2 <- bird_data[check2,]
  
```

# Similarity over all the different variables

## Perceptual Data

### Add AVONET data

```{r}

#prepare avonet dataset

AVONET<-read.csv(here("preprocessing/data/AVONET2_eBird.csv")) %>% 
  rename(
    c(ScientificName="Species2",
      Family=Family2,
      Order=Order2,
      BeakLengthCulmen="Beak.Length_Culmen",
      BeakLengthNares="Beak.Length_Nares",
      BeakWidth="Beak.Width",
      BeakDepth="Beak.Depth",
      TarsusLength="Tarsus.Length",
      WingLength="Wing.Length",
      KippsDistance="Kipps.Distance",
      Secondary="Secondary1",
      HandWingIndex="Hand.Wing.Index",
      TailLength="Tail.Length",
      HabitatDensity="Habitat.Density",
      TrophicLevel="Trophic.Level",
      TrophicNiche="Trophic.Niche",
      Lifestyle="Primary.Lifestyle")) %>% 
  select(c(ScientificName,
           Family,
           Order,
           BeakLengthCulmen,
           BeakLengthNares,
           BeakWidth,
           BeakDepth,
           TarsusLength,
           WingLength,
           KippsDistance,
           Secondary,
           HandWingIndex,
           TailLength,
           Mass,
           Habitat,
           HabitatDensity,
           Migration,
           TrophicLevel,
           TrophicNiche,
           Lifestyle))

bird_data <- left_join(bird_data, 
                       AVONET %>%
                    select(ScientificName,
                           BeakLengthCulmen:Mass), 
                    by = c(
                      "ScientificName2021" = "ScientificName")) 

```

### Calculate Perceptual Similarity & Isolation

```{r}

pairwise_list_per <- list()
per_dist_matrices <- list()
bird_data$IsoPer <- NA
bird_data$zIsoPer <- NA

languages <- c("Anindilyakwa",
               "Montagnais",
               "Rangi",
               "Saami",
               "Tlingit",
               "Tobelo",
               "Tzeltal",
               "Zapotec")

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    select(CanonicalName,BeakLengthCulmen:Mass) %>% 
    filter(complete.cases(.[sapply(., is.numeric)])) %>%
    mutate_if(is.numeric, log) %>% # log transform the skewed variables
    mutate(HandWingIndex = exp(HandWingIndex)) %>% # un-transform the variable that isn't skewed
    mutate_if(is.numeric, ~scale(.)[,1]) #This fixes that pesky problem where it changes the column name when it scales
  
  data_for_indices <- bird_data %>% 
    filter(Language == language)
  
  # Make a matrix with euclidean distances between all birds based on their AVONET traits
  matrices <- language_data  %>%
    select(c(BeakLengthCulmen:Mass)) %>% 
    data.matrix() %>% 
    pdist(metric = "euclidean", p = 2) %>% 
    as.data.frame() 
  
  rownames(matrices) <- language_data$CanonicalName
  colnames(matrices) <- language_data$CanonicalName
  
  # Put perceptual isolation values back into main dataframe (The distance between a bird and its closest neighbour) - not sure if I'm using these but I might as well put them in now in case I want them later

  iso <- MinVal(matrices) %>% 
    mutate(pure_names = gsub("\\..*", "", Names))
  
  ziso <- iso %>% mutate(pure_names =  gsub("\\..*", "", Names))%>% 
    distinct(pure_names, .keep_all = TRUE) %>% 
    mutate(zIsoPer = round(scale(Values)[,1],2)) %>% 
    select(-Names)
  
  iso <- left_join(iso, ziso) %>% 
    mutate(IsoPer = round(Values,2),
           CanonicalName = Names) %>% 
    select(CanonicalName, IsoPer, zIsoPer)
    
  language_data <- left_join(language_data,iso, by = "CanonicalName")

    # language_data$IsoPer <- round(iso$Values,2)
    # language_data <- language_data %>% 
    #   mutate(zIsoPer = round(scale(IsoPer)[,1],2))
    
    bird_data$IsoPer[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$IsoPer
    bird_data$zIsoPer[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$zIsoPer

    colnames(matrices) <- language_data$CanonicalName
    rownames(matrices) <- language_data$CanonicalName
    
    # Put the values into a long dataframe
    per_dist_list <- data.frame(t(combn(names(matrices),2)),
                              DistPer = t(matrices)[lower.tri(matrices)]) %>% 
      mutate(Species1 = pmin(X1, X2),
             Species2 = pmax(X1, X2),
             Species1FG = data_for_indices$IndexFG[match(Species1, language_data$CanonicalName)],
             Species2FG = data_for_indices$IndexFG[match(Species2, language_data$CanonicalName)],
             SameName = ifelse(Species1FG == Species2FG, TRUE, FALSE)) %>% 
      select(-c(X1, X2, Species1FG, Species2FG))
    
    pairwise_list_per[[language]] <- per_dist_list
    per_dist_matrices[[language]] <- matrices
    
    if (language=="Montagnais") {
      
      write.csv(matrices, here("data/per-matrix_Innu.csv"))
    } else {
    
    write.csv(matrices, here(paste0("data/per-matrix_", language, ".csv")))
      
    }
}

```

### Get tSNE coordinates

```{r}

bird_data <- bird_data %>% mutate(tSNE1per = NA, tSNE2per = NA)

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    select(-contains("tSNE"))
  
  # I need to take out the duplicate rows and it's kind of tricky because the canonical names are given suffixes based on duplicates over the entire bird dataset. So I am going to remove the suffixes from the rownames, remove duplicates, then transpose the dataframe and do it again.
  
  tsne_data <- per_dist_matrices[[language]] %>% 
    rownames_to_column("name") %>% 
    mutate(name = str_remove(name, "\\.\\d")) %>% 
    filter(!duplicated(name)) %>% 
    column_to_rownames("name") %>% 
    t() %>% 
    as.data.frame() %>% 
    rownames_to_column("name") %>% 
    mutate(name = str_remove(name, "\\.\\d")) %>% 
    filter(!duplicated(name)) %>% 
    column_to_rownames("name")
    
  # Perplexity should not be greater than 3 * perplexity < nrow(X) - 1 according to documentation for Rtsne package. So I'm just reverse-engineering the perplexity parameter
  set.seed(29)
  tsne_fit <- Rtsne(tsne_data,
                    perplexity = floor(nrow(tsne_data) - 1) / 3)
  
  tsne_layout <- tsne_fit$Y %>% 
    as.data.frame() %>% 
    rename(tSNE1per = "V1",
           tSNE2per = "V2")
  
  tsne_data <- data.frame(tsne_layout,
                          ScientificName2023 = rownames(tsne_data))
  
  language_data <- left_join(language_data, tsne_data, relationship = "many-to-one", by = "ScientificName2023")
  
  bird_data$tSNE1per[match(
    language_data$CanonicalName,
    bird_data$CanonicalName)] <- language_data$tSNE1per
  
  bird_data$tSNE2per[match(
    language_data$CanonicalName,
    bird_data$CanonicalName)] <- language_data$tSNE2per
}

```

### Might just calculate some variances while I'm here

```{r}

bird_data$VarPer <- NA

for (language in languages) {
  
  # Make a dataframe of centroids for each FG group
  
  language_data <- bird_data %>% 
    filter(Language == language,
           n>1) %>% 
    mutate_at(vars(BeakLengthCulmen:Mass), log) %>%  # make all the positively skewed ones normal
    mutate(HandWingIndex = exp(HandWingIndex)) %>% # un-transform the variable that isn't skewed
    mutate_at(vars(BeakLengthCulmen:Mass), ~scale(.)[,1])  # scale them all

  centroids <- language_data %>% 
    pivot_longer(cols = BeakLengthCulmen:Mass,
                 names_to = "Feature",
                 values_to = "Values") %>% 
    
    # Calculate the centroid for each trait for each FG and save it as a new bird 
    
    group_by(Feature, IndexFG) %>% 
    summarise(Centroid = mean(Values)) %>% 
    pivot_wider(names_from = "Feature",
                values_from = "Centroid") %>% 
    mutate(IndexFG = paste0(IndexFG, ".cen"))
  
  # Put the centroid birds back into language_data
  
  birds_cens <- language_data %>% 
    filter(n > 1) %>% 
    select(IndexFG, BeakLengthCulmen:Mass) %>% 
    rbind(centroids)
  
  # Make a distance matrix of all birds and centroids
  
  matrix <- birds_cens %>% 
    select(-c(IndexFG)) %>% 
    pdist(metric = "euclidean") %>% 
    as.data.frame()
  
  colnames(matrix) <- birds_cens$IndexFG
  matrix$IndexFG <- birds_cens$IndexFG
  
  # Turn it into a long dataframe and keep only the distances between birds and their own centroids
  
  variance <- matrix %>% 
    pivot_longer(cols = -contains("IndexFG"),
                 names_to = "Group",
                 values_to = "Dist") %>% 
  filter(grepl("\\.cen$", Group) & !grepl("\\.cen$", IndexFG)) %>% 
    mutate(Group = gsub("\\.cen$", "", Group)) %>% 
    filter(IndexFG == Group) %>% 
    group_by(Group)   %>%
    summarise(Variance = mean(Dist))
  
  bird_data <- bird_data %>%
    mutate(VarPer =
             ifelse(IndexFG %in% variance$Group,
                    variance$Variance[match(IndexFG, variance$Group)],
                    VarPer),
           # Give monotypic birds a variance of 0
           VarPer = 
             ifelse(n == 1, 0, VarPer))
  }

```

## Phylogenetic Data

### Process Phylogenetic Data

Phylogeny subsets (250 trees)

anindilyakwa - tree-pruner-ee9aeff3-6c7d-4b87-aa87-7d38474286fd 
montagnais - tree-pruner-65a824c3-96d0-4874-a14e-5c16cef86174 
rangi - tree-pruner-a6edba71-4114-4d77-a124-33a41b5e040a
saami - tree-pruner-9a0ae6e0-7769-4035-b94d-6990fd7a0130
tlingit - tree-pruner-d5b77082-3468-46aa-9dfc-1ea52bd48562
tobelo - tree-pruner-9c9972ae-a991-47a3-9d35-01692750c6e4 
tzeltal - tree-pruner-c45e9238-3a1a-4eaa-a980-b810c0b1954c 
zapotec - tree-pruner-1546ad01-71c3-4e2b-a201-11f91ccccd83 

```{r}

phy_dist_matrices <- list()
nexuses <- list()

for (language in languages) {
  
  print(language)
  
  # import nexuses
  location <- here(paste0("preprocessing/data/nexus_", str_to_lower(language), ".nex")) 
  
  nexus <- read.nexus(location)
  
   # turn it into a bunch of matrices
   phylo_matrix_full <- lapply(nexus, function(tree) cophenetic(tree))
  
   # average out the matrices
   phylo_matrix <- Reduce("+", phylo_matrix_full) / length(phylo_matrix_full)
  
   # Make the names in the matrices match the language datasets
   rownames(phylo_matrix) <- gsub("_", " ", rownames(phylo_matrix))
   colnames(phylo_matrix) <- gsub("_", " ", colnames(phylo_matrix))
  
   # Process tree so it can be visualised later:
  
   # create a consensus tree but preserve edge lengths
   phylo_tree <- consensus.edges(nexus, method = "mean.edge")

   # Drop the last edge length, which doesn't correspond with an edge and starts messing stuff up later
   phylo_tree$edge.length <- phylo_tree$edge.length[-length(phylo_tree$edge.length)]

   phylo_tree <- chronos(phylo_tree, lambda = 0)

   # Write it so it can be imported directly
   
   if (language == "Montagnais") {
     
     write.nexus(phylo_tree, file = here("data/phylo-tree_Innu.nex")) 
   } else {

   write.nexus(phylo_tree, file = here("data/", paste0("phylo-tree_", language, ".nex")))
     
   } 
   
    phy_dist_matrices[[language]] <- phylo_matrix
   nexuses[[language]] <- nexus
  
}

```


```{r eval=FALSE}

montagnais_wrong_names <- nexuses$montagnais$tree_2141$tip.label %>% 
  as.data.frame() %>% 
  rename(ScientificNameV3 = ".") %>% 
  mutate(ScientificNameV3 = gsub("_"," ", ScientificNameV3))

in_data_not_nexus <- setdiff(montagnais$ScientificNameV3, montagnais_wrong_names$ScientificNameV3)

in_nexus_not_data <- setdiff(montagnais_wrong_names$ScientificNameV3,
                             montagnais$ScientificNameV3)

```

Lengths don't line up:
Anindilyakwa - 1, because of the double Egretta sacra
Montagnais - 1, because of double Somateria mollissima
Rangi - 9 because legitimate duplicates
Saami - 1 because of double Larus argentatus
Tobelo - 8 because legitimate duplicates
Tzeltal - 1 because Aratinga canicularis
Tlingit is good
Zapotec is good

### Calculate Phylogenetic Similarity & Isolation

```{r}

pairwise_list_phy <- list()
matrix2 <- list()
bird_data$IsoPhy <- NA
bird_data$zIsoPhy <- NA
check <- list()
check_vals <- list()
check_iso <- list()

for (language in languages) {
  
  # Oops I messed up by sometimes having the languages all lowercase 
  ## Yeah, thanks. This is Future Zoë going and fixing it all.
  language_data <- bird_data %>% 
    filter(Language == language)
  
  matrix <- as.data.frame(phy_dist_matrices[[language]]) %>% 
    rownames_to_column(var = "ScientificNameV3")
  
  # Deal with duplicate birds
  
  duplicates <- language_data[duplicated(language_data$ScientificNameV3),]
  
  # Make duplicate rows of the duplicate birds (This will work no matter how many times the bird is duplicated)
  
  rows_add <- duplicates %>% 
    select(ScientificNameV3) %>% 
    left_join(matrix)
  
  matrix <- rbind(matrix, rows_add)
  
  duplicate_names <- as.array(duplicates$ScientificNameV3)
  
  # Make duplicate columns of the duplicated birds. This won't work if there are more than two of any bird
  
  cols_add <- matrix %>% 
    select(all_of(duplicate_names)) %>% 
    rename_all(~paste0(.,".1"))
  
  # This will add in columns for any bird in the dataset three times
  
  double_duplicates <- duplicate_names[duplicated(duplicate_names)]
  
  if(length(double_duplicates) != 0) {
    
    cols_add <- matrix %>% 
      select(all_of(double_duplicates)) %>% 
      rename_all(~paste0(.,".2")) %>% 
      bind_cols(cols_add)
  }
  
  merge_data <- language_data %>% 
    mutate(ScientificNameV3 = make.unique(ScientificNameV3)) %>% 
    select(ScientificNameV3,
           CanonicalName)
  
  # Pop the unique canonical names into the matrix
  matrix <- bind_cols(matrix, cols_add) %>% 
    # select(order(colnames(.))) %>% 
    rename_all(~ ifelse(. %in% merge_data$ScientificNameV3, merge_data$CanonicalName[match(.,merge_data$ScientificNameV3)],.)) %>% 
    mutate(ScientificNameV3 = make.unique(ScientificNameV3))  %>% 
   left_join(merge_data, by = "ScientificNameV3") %>% 
   column_to_rownames(var = "CanonicalName") %>% 
   select(-ScientificNameV3)
   
  # Make sure the columns and rows are ordered the same
   sort_data <- rownames(matrix)
   matrix <- matrix[sort_data,sort_data]
   
     # write it so it can be downloaded again directly
   
   if (language=="Montagnais") {
     
     write.csv(matrix, here("data/phylo-matrix_Innu.csv"))
     
   } else {
     
   write.csv(matrix, here("data/", paste0("phylo-matrix_", language, ".csv")))
     
   }
   
   check[[language]] <- matrix
   check_rows <- rownames(matrix) %>% as.data.frame() %>% rename(rows = ".")
   check_cols <- colnames(matrix) %>% as.data.frame() %>% rename(cols = ".")
   check_vals[[language]] <- cbind(check_rows,check_cols)  %>% 
      mutate(Same = ifelse(rows==cols,TRUE,FALSE))
   
   iso <- MinVal(matrix)   %>%
     mutate(Values = as.numeric(Values))
   
   check_iso[[language]] <- iso
   
   ziso <- iso %>% mutate(pure_names =gsub("\\..*", "", Names)) %>% 
     distinct(pure_names, .keep_all = TRUE) %>% 
     mutate(zIsoPhy = round(scale(Values)[,1],2)) %>% 
     select(-c(Names, Values))
  
  iso <- iso %>% 
    mutate(pure_names = gsub("\\..*", "", Names)) %>% 
    left_join(ziso, by = "pure_names") %>% 
    mutate(IsoPhy = round(Values,2),
           CanonicalName = Names) %>% 
    select(CanonicalName, IsoPhy, zIsoPhy)
   
   language_data$IsoPhy[match(iso$CanonicalName,
                              language_data$CanonicalName)] <- 
     iso$IsoPhy
   
      language_data$zIsoPhy[match(iso$CanonicalName,
                              language_data$CanonicalName)] <- 
     iso$zIsoPhy
   

    
    bird_data$IsoPhy[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$IsoPhy
    
    bird_data$zIsoPhy[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$zIsoPhy
  
  #Turn the matrix into a pairwise list
  phy_dist_list <- data.frame(t(combn(rownames(matrix),2)),
                              DistPhy = t(matrix)[lower.tri(matrix)])
  
  pairwise_list_phy[[language]] <- phy_dist_list %>% 
    mutate(Species1 = pmin(X2, X1),
           Species2 = pmax(X2, X1)) %>% 
    select(-c(X2, X1))
  
}

```

## Taxonomic Similarity 
### 2023

```{r}

pairwise_list_23 <- list()
bird_data$Iso23 <- NA
bird_data$zIso23 <- NA

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language)
  
  to_compare <- c("Order2023", 
                  "Family2023", 
                  "Genus2023", 
                  "ScientificName2023")
  
  similarity_2023 <- TaxSimCalc(language_data, to_compare)
  
  matrix23 <- similarity_2023$matrix %>% column_to_rownames("Species1")
  
  iso23 <- MinVal(matrix23) %>%
     mutate(Values = as.numeric(Values))
   
     ziso <- iso23 %>% mutate(pure_names =  gsub("\\..*", "", Names)) %>% 
    distinct(pure_names, .keep_all = TRUE) %>% 
    mutate(zIso23 = round(scale(Values)[,1],2)) %>% 
    select(-c(Names, Values))
  
  iso23 <- iso23 %>% mutate(pure_names = gsub("\\..*", "", Names)) %>% 
    left_join( ziso, by = "pure_names") %>% 
    mutate(Iso23 = round(Values,2),
           CanonicalName = Names) %>% 
    select(CanonicalName, Iso23, zIso23)
  
  language_data$Iso23[
    match(language_data$CanonicalName,iso23$CanonicalName)] <- 
    iso23$Iso23
  language_data$zIso23[
    match(language_data$CanonicalName,iso23$CanonicalName)] <- 
    iso23$zIso23
  
  
    
  bird_data$Iso23[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$Iso23
  
  bird_data$zIso23[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$zIso23  

  pairwise_list_23[[language]] <- similarity_2023$pairs %>% 
    rename(Dist23 = Dist)
  
}

```

### 1974

```{r}

pairwise_list_74 <- list()
bird_data$Iso74 <- NA
bird_data$zIso74 <- NA

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    select(-c(Iso74, zIso74))
  
  to_compare <- c("Order1974", 
                  "Family1974", 
                  "Genus1974", 
                  "ScientificName1974")
  
  similarity_1974 <- TaxSimCalc(language_data, to_compare)
  
  matrix74 <- similarity_1974$matrix %>% column_to_rownames("Species1")
  
  iso74 <- MinVal(matrix74) %>%
     mutate(Values = as.numeric(Values))
   
     ziso <- iso74 %>% mutate(pure_names =  gsub("\\..*", "", Names))%>% 
    distinct(pure_names, .keep_all = TRUE) %>% 
    mutate(zIso74 = round(scale(Values)[,1],2)) %>% 
    select(-c(Names, Values))
  
  iso74 <- iso74 %>% 
    mutate(pure_names = gsub("\\..*", "", Names)) %>% 
    left_join(ziso, by = "pure_names") %>% 
    mutate(Iso74 = round(Values,2),
           CanonicalName = Names) %>% 
    select(CanonicalName, Iso74, zIso74)
    
    language_data <- left_join(language_data, iso74, by = "CanonicalName")
  
  # language_data$Iso74[match(language_data$CanonicalName,iso74$CanonicalName)] <- 
  #   iso74$Iso74
  # language_data$zIso74[match(language_data$CanonicalName,iso74$CanonicalName)] <- 
  #   iso74$zIso74

  bird_data$Iso74[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$Iso74
  
  bird_data$zIso74[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$zIso74
  
  pairwise_list_74[[language]] <- similarity_1974$pairs %>% 
    rename(Dist74 = Dist)
}

```

## Colour Data

I'm going to do with the colours the same as what Andrea Santangeli et al. did and just keep black, white, yellow, blue, red, and green, and use the sum of the light and dark versions (p. 20 of paper)


```{r}

colour_data <- read.csv(here("preprocessing/data/colour_data.csv")) %>%  
  rename(ScientificNameV3 = BirdTREE,
         ScientificName2021 = sciName_ebird2021,
         cDiversity = n.loci,
         cElaboration = elaboration,
         cBlack = black,
         cWhite = white,
         cYellow = yellow) %>% 
  mutate(cRed = redD + redL,
         cBlue = blueD + blueL,
         cGreen = greenD + greenL) %>% 
filter(!is.na(cElaboration)) %>%
  arrange(desc(cElaboration)) %>% 
  select(ScientificName2021,
         ScientificNameV3,
         cBlack,
         cWhite,
         cYellow,
         cBlue,
         cRed,
         cGreen,
         cElaboration,
         cDiversity,
         sex,
         sciname_ebird2019) %>% 
  # Keep the most colourful set of observations for each bird
  distinct(ScientificName2021, .keep_all = TRUE)

# They haven't split apart two gull species that ebird 2021 differentiates, I'll just double that guy's row for them both.
duplicate_row <- colour_data[which(colour_data$ScientificName2021 == "Larus canus"),]
duplicate_row$ScientificName2021 <- "Larus brachyrhynchus"
colour_data <- colour_data %>% rbind(duplicate_row)


colours <- bird_data %>% 
  left_join(colour_data %>% select(-c(sex,sciname_ebird2019,ScientificNameV3)), 
            by = "ScientificName2021",
            relationship = "many-to-one") 

bird_data <- colours %>% 
  filter(is.na(cElaboration)) %>% 
  select(!c(cBlack:cDiversity)) %>% 
  left_join(colour_data %>% select(-c(sex,sciname_ebird2019,ScientificName2021)), 
            by = "ScientificNameV3",
            relationship = "many-to-one") %>% 
  rbind(colours %>% filter(!is.na(cElaboration)))

```

### Colour similarity

```{r}

pairwise_list_col <- list()

bird_data$IsoCol <- NA
bird_data$zIsoCol <- NA
bird_data$zcElaboration <- NA
bird_data$zcDiversity <- NA

for (language in languages) { 
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    select(-c(IsoCol, zIsoCol)) %>% 
    select(CanonicalName,cBlack:cGreen) %>% 
    filter(complete.cases(.[sapply(., is.numeric)])) %>%
    mutate_if(is.numeric, ~scale(.)[,1]) #This fixes that pesky problem where it changes the column name when it scales
  
  data_for_indices <- bird_data %>% 
    filter(Language == language)
  
  # Make a matrix with euclidean distances between all birds based on their colours
  matrices <- language_data  %>%
    select(c(cBlack:cGreen)) %>% 
    data.matrix() %>% 
    pdist(metric = "euclidean", p = 2) %>% 
    as.data.frame()
  
  colnames(matrices) <- language_data$CanonicalName
  rownames(matrices) <- language_data$CanonicalName
  
  # Put perceptual isolation values back into main dataframe (The distance between a bird and its closest neighbour) - not sure if I'm using these but I might as well put them in now in case I want them later
  
  iso <- MinVal(matrices) %>% 
    mutate(pure_names = gsub("\\..*", "", Names))
  
  ziso <- iso %>% 
    distinct(pure_names, .keep_all = TRUE) %>% 
    mutate(zIsoCol = round(scale(Values)[,1],2)) %>% 
    select(-c(Names, Values))
  
  iso <- left_join(iso, ziso, by = "pure_names") %>% 
    mutate(IsoCol = round(Values,2)) %>% 
    rename(CanonicalName = Names) %>% 
    select(CanonicalName, IsoCol, zIsoCol)

  language_data <- left_join(language_data, iso, by = "CanonicalName")

    bird_data$IsoCol[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$IsoCol
    bird_data$zIsoCol[match(language_data$CanonicalName, 
                           bird_data$CanonicalName)] <- 
      language_data$zIsoCol
    
    colnames(matrices) <- language_data$CanonicalName
    rownames(matrices) <- language_data$CanonicalName
    
    # Put the values into a long dataframe
    col_dist_list <- data.frame(t(combn(names(matrices),2)),
                              DistCol = t(matrices)[lower.tri(matrices)]) %>% 
      mutate(Species1 = pmin(X1, X2),
             Species2 = pmax(X1, X2)) %>% 
    select(-c(X1, X2))
    

    pairwise_list_col[[language]] <- col_dist_list
    
}

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    mutate(zcDiversity = round(scale(cDiversity)[,1],2),
           zcElaboration = round(scale(cElaboration)[,1],2))
  
  # language_data <- language_data[complete.cases(language_data),]
  
  #This feels like a bit of a cowboy way to get the scaled Diversity and Elaboration scores back into the dataset, but it works
  
bird_data <- bird_data %>% 
  mutate(zcDiversity = ifelse(CanonicalName %in% language_data$CanonicalName,
                              language_data$zcDiversity[match(
                                bird_data$CanonicalName,
                                language_data$CanonicalName)],
                              zcDiversity))
  bird_data <- bird_data %>% 
  mutate(zcElaboration = ifelse(CanonicalName %in% language_data$CanonicalName,
                              language_data$zcElaboration[
                                match(bird_data$CanonicalName,
                                      language_data$CanonicalName)],
                              zcElaboration))
  
}

```

## Smoosh all of the similarity lists together

```{r}

bird_similarity <- list()

for (language in languages) {
  
  per <- pairwise_list_per[[language]]
  phy <- pairwise_list_phy[[language]]
  tax74 <- pairwise_list_74[[language]]
  tax23 <- pairwise_list_23[[language]]
  col <- pairwise_list_col[[language]]
  
  data <- list(per,
               phy,
               tax74,
               tax23,
               col)
  
  merged_data <- Reduce(function(x, y)
    merge(x, y,
          by = c("Species1", "Species2"),
          all = TRUE),
    data) %>% 
    select(-c(starts_with("Scale")))
  
  bird_similarity[[language]] <- merged_data
  
  if (language=="Montagnais") {
    
    write.csv(merged_data, here("data/similarity_Innu.csv"))
    
  } else {
  
  write.csv(merged_data, here(paste0("data/similarity_", language, ".csv")))
    
  }
}

bird_similarity[["Innu"]] <- bird_similarity[["Montagnais"]]

```

## Calculate variances for phylogenetic

### Variance in phylogenetic space

```{r}

bird_data$VarPhy <- NA

for (language in languages) {
  
  language_data <- bird_data %>% 
    filter(Language == language) %>% 
    select(-VarPhy)
    
  phy_var_data <- bird_similarity[[language]] %>% 
    filter(SameName == TRUE)  %>% 
    left_join(language_data %>% select(CanonicalName, IndexFG), 
              by = c("Species1" = "CanonicalName"),
              relationship = "many-to-many") %>% 
    group_by(IndexFG) %>% 
    mutate(VarPhy = mean(DistPhy))
  
  language_data <- language_data %>% 
    left_join(phy_var_data %>% select(VarPhy, IndexFG),
              relationship = "many-to-many") %>% 
    mutate(VarPhy = ifelse(n==1, 0, VarPhy))
  
  bird_data <- bird_data %>% 
    mutate(VarPhy = ifelse(IndexFG %in% language_data$IndexFG,
                           language_data$VarPhy[match(IndexFG, 
                                                      language_data$IndexFG)], 
                           VarPhy))
  
  # bird_data$VarPhy <- language_data$VarPhy[match(bird_data$IndexFG, language_data$IndexFG)]
  
}

```

# Taxonomic Trees

This makes trees out of the levels of the traditional taxonomies

## 1974 Trees

```{r}

tree_74 <- list()

for (language in languages) {
  
  language_data <- bird_data %>% 
    ungroup() %>% 
    filter(Language == language)
  
    # This is to keep the families in a particular order if I want to visualise the tree later
family_order_74 <- orders_74 %>% 
  select(Family1974) %>% 
  filter(Family1974 %in% language_data$Family1974) %>% 
  mutate(Family1974 = as.factor(Family1974))
  
  language_data <- language_data %>% 
    mutate(Class = "Aves") %>% 
    select(ScientificName1974,
           Genus1974,
           Family1974,
           Order1974,
           Class) %>% 
    distinct(ScientificName1974, .keep_all = TRUE) %>% 
    mutate_all(as.factor) %>% 
     arrange(match(Family1974, family_order_74))
   
# Make the data into a tree structure based on the levels of the taxonomy
  
  tree <- as.phylo(~Class/Order1974/Family1974/Genus1974/ScientificName1974,
                      data = language_data) %>% 
    compute.brlen(.25) %>% # Add branch lengths so they can be used later
    as.ultrametric()
  
  tree_74[[language]] <- tree
  
  if (language == "Montagnais"){
    
         write.nexus(tree, 
               file = here(paste0("data/tree_74_Innu.nex")))
  } else {
  
     write.nexus(tree, 
               file = here(paste0("data/tree_74_", language, ".nex")))
    
  }
  }

```

## 2023 Trees

```{r}

tree_23 <- list()

for (language in languages) {

    language_data <- bird_data %>% 
    ungroup() %>% 
    filter(Language == language)
  
  family_order_23 <- clements_2023 %>% 
  select(Family2023) %>% 
  filter(Family2023 %in% language_data$Family2023) %>% 
  mutate(Family2023 = as.factor(Family2023))
  
  language_data <- language_data %>% 
    mutate(Class = "Aves") %>%
    group_by(ScientificName2023) %>% 
    # mutate(Subspecies = case_when(
    #   n_distinct(IndexFG) > 1 ~ make.unique(ScientificName2023), TRUE ~ ScientificName2023
    # )) %>% 
    ungroup() %>% 
    select(ScientificName2023,
           Genus2023,
           Family2023,
           Order2023,
           Class) %>% 
    distinct(ScientificName2023, .keep_all = TRUE) %>% 
    mutate_all(as.factor) %>% 
    arrange(match(Family2023, family_order_23))
  
  tree <- as.phylo(~Class/Order2023/Family2023/Genus2023/ScientificName2023,
                   data = language_data) %>% 
    compute.brlen(1) 
  # %>% 
  #   as.ultrametric()
  
  tree_23[[language]] <- tree
  
  if(language == "Montagnais") {
    
    write.nexus(tree, file = here(paste0("data/tree_23_Innu.nex")))
    
  } else {
  
  write.nexus(tree, file = here(paste0("data/tree_23_", language  , ".nex")))
    
  }
}

```

# Export the data

```{r}

# Change the old name for Innu

bird_data <- bird_data %>% 
  mutate(Language = ifelse(Language == "Montagnais", "Innu", Language))

write.csv(bird_data,here("data/bird_data.csv")) 

```

# Summary

# For Isolation Data

```{r}

# Summary for all Isolation data

# Select columns that start with "iso"
iso_columns <- grep("^Iso", names(bird_data), value = TRUE)

# Group by "language" and calculate summary statistics
summary_iso <- bird_data %>%
  group_by(Language) %>%
  add_count(name = "n") %>% 
  summarise(across(all_of(iso_columns), list(
    mean = ~mean(., na.rm = TRUE),
    sd = ~sd(., na.rm = TRUE),
    min = ~min(., na.rm = TRUE),
    max = ~max(., na.rm = TRUE)
  ), .names = "{col}_{fn}"),
  n = first(n)) %>% 
  mutate_if(is.numeric, ~round(.,2))

# Take out outliers and duplicates and recalculate stats

duplicates <- bird_data$ScientificName2023[duplicated(bird_data$ScientificName2023)]
duplicates <- bird_data %>% filter(ScientificName2023 %in% duplicates) %>% 
  group_by(Language, ScientificName2023) %>% 
  mutate(number = n()) %>% 
  filter(number != 1)

bird_data2 <- bird_data %>% 
  filter(!CanonicalName %in% duplicates$CanonicalName) %>% 
  filter(ScientificName2023 != "Dromaius novaehollandiae") %>% 
  filter(ScientificName2023 != "Struthio camelus")

summary_iso2 <- bird_data2 %>% 
  group_by(Language) %>%
   add_count(name = "n") %>% 
  summarise(across(all_of(iso_columns), list(
    mean = ~mean(., na.rm = TRUE),
    sd = ~sd(., na.rm = TRUE),
    min = ~min(., na.rm = TRUE),
    max = ~max(., na.rm = TRUE)
  ), .names = "{col}_{fn}"),
  n = first(n)) %>% 
  mutate_if(is.numeric, ~round(.,2)) %>% 
  mutate(Language = paste0(Language, "_filtered"))

cols <- c(summary_iso$Language, summary_iso2$Language)

summary_iso_full <- rbind(summary_iso, summary_iso2) %>% 
  as.data.frame() %>% 
  t() %>% 
  as.data.frame()

colnames(summary_iso_full) <- cols

summary_iso_full <- summary_iso_full[,order(names(summary_iso_full))]

 # write.csv(summaryfull, here("output/SummaryTable.csv"))

```

# For Distance Data

```{r}

summary_dist <- data.frame(matrix(nrow = 20))

for (language in languages) {
  
  language_data <- bird_similarity[[language]] 
  
  # Calculate summary statistics
  summary_lang <- language_data %>%
  summarise(across(where(is.numeric), list(
    mean = ~mean(., na.rm = TRUE),
    sd = ~sd(., na.rm = TRUE),
    min = ~min(., na.rm = TRUE),
    max = ~max(., na.rm = TRUE)), 
    .names = "{col}_{fn}")) %>% 
  mutate_if(is.numeric, ~round(.,2)) %>% 
    t() %>% 
    as.data.frame() %>% 
    rename_with(~ language, .cols = 1)
  
  summary_dist <- bind_cols(summary_dist, summary_lang)
  
  summary_lang2 <- language_data %>% 
    filter(Species1 %in% bird_data2$CanonicalName & 
             Species2 %in% bird_data$CanonicalName) %>% 
    summarise(across(where(is.numeric), list(
      mean = ~mean(., na.rm = TRUE),
      sd = ~sd(., na.rm = TRUE),
      min = ~min(., na.rm = TRUE),
      max = ~max(., na.rm = TRUE)),
      .names = "{col}_{fn}")) %>% 
    mutate_if(is.numeric, ~round(.,2)) %>% 
    t() %>% 
    as.data.frame() %>% 
    rename_with(~ paste0(language, "_filtered"))
  
 
    summary_dist <- bind_cols(summary_dist, summary_lang2)
  
}

summary_dist <- summary_dist %>% 
  rename_with(~ gsub("Montagnais", "Innu", .))

summary_full <- summary_dist %>% 
  select(-1) %>% 
  rename_all(~str_to_title(.)) %>% 
  rbind(summary_iso_full) %>% 
  subset(!rownames(.) == "Language")

fil_col <- grep("_filtered$", colnames(summary_full), value = TRUE)

for (col in fil_col) {
  
  non_fil_col <- sub("_filtered$", "", col)
  
  if (all(summary_full[[col]] == summary_full[[non_fil_col]], na.rm = TRUE)) {
    summary_full[[col]] <- NULL
  }
  
}

 write.csv(summary_full, here("output/SummaryTable.csv"))

```
