---
title: "Analysis 1 with cross-validation: which measures predict name-sharing?"
output:
  html_document:
    toc: true
---

This notebook reports the cross-validated version of Analysis 1. It reads the
cached results written by `cv_analyses/run_cv.R` and does no model fitting
itself. See `cv_analyses/README.md` for the design.

```{r setup, message = FALSE}
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)

out_dir <- here("cv_analyses", "output")
source(here("cv_analyses", "cv_functions.R"))

lang_order <- c(languages, "All")
model_order <- names(models)
contrast_order <- names(contrasts)

insample <- read.csv(file.path(out_dir, "in_sample.csv"))
cv <- readRDS(file.path(out_dir, "cv_species.rds"))
lolo <- readRDS(file.path(out_dir, "cv_lolo.rds"))
run_info <- readLines(file.path(out_dir, "run_info.txt"))

cv_sum <- summarise_cv(cv)
cv_con <- summarise_contrasts(cv)
lolo_con <- bind_rows(summarise_lolo_contrasts(lolo, "ll_pop"),
                      summarise_lolo_contrasts(lolo, "ll_recal"))

n_reps <- n_distinct(cv$rep)
```

Settings: `r paste(run_info[1:3], collapse = ", ")`.

# 1. Species held out, within each language

For each language, species were split at random into 10 folds. Each model was
fitted to the pairs in which neither species was in the held-out fold, and
scored on the pairs in which at least one was. Every pair is scored once, so
the summed held-out log-likelihood (LL) is on the same scale as the in-sample
LL. This was repeated with `r n_reps` random fold splits. The "All" column is
a pooled model with a separate intercept per language and common slopes.

## Held-out deviance compared with in-sample BIC

Each cell shows the held-out deviance difference from the best model,
ΔDev = 2 × (LL~best~ − LL~model~), averaged over splits. Below it, in
brackets, is the in-sample ΔBIC from Table 2 of the paper. Both are 0 for the
best model, and both are on the deviance scale. ΔBIC adds a penalty for each
extra parameter, while ΔDev penalises overfitting by testing on held-out pairs.

```{r table-ddev}
ddev_tab <- cv_sum %>%
  select(language, model, ddev_mean, ddev_sd, pct_best) %>%
  left_join(insample %>% select(language, model, dbic),
            by = c("language", "model")) %>%
  mutate(language = factor(language, lang_order),
         model = factor(model, model_order)) %>%
  arrange(model, language)

write.csv(ddev_tab, file.path(out_dir, "table_ddev.csv"), row.names = FALSE)

fmt <- function(x) formatC(x, format = "f", digits = 1)

ddev_wide <- ddev_tab %>%
  mutate(cell = paste0(fmt(ddev_mean), " (", fmt(dbic), ")")) %>%
  select(model, language, cell) %>%
  pivot_wider(names_from = language, values_from = cell)

kable(ddev_wide, align = c("l", rep("r", length(lang_order))),
      caption = "Held-out ΔDev, with in-sample ΔBIC in brackets") %>%
  kable_styling(full_width = FALSE)

latex_ddev <- ddev_wide %>%
  kable(format = "latex", booktabs = TRUE, linesep = "",
        align = c("l", rep("c", length(lang_order))),
        caption = paste("Cross-validated model comparison for name-sharing.",
                        "Cells show the held-out deviance difference from the",
                        "best model (mean over", n_reps, "random 10-fold",
                        "species splits), with the in-sample $\\Delta$BIC in",
                        "brackets."),
        label = "table:cvLR", escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
writeLines(latex_ddev, file.path(out_dir, "table_ddev.tex"))
```

How often each model came out best, as a percentage of splits:

```{r table-pct-best}
ddev_tab %>%
  select(model, language, pct_best) %>%
  pivot_wider(names_from = language, values_from = pct_best) %>%
  kable(align = c("l", rep("r", length(lang_order)))) %>%
  kable_styling(full_width = FALSE)
```

AUC (mean over splits) for each model's held-out predictions:

```{r table-auc}
cv_sum %>%
  mutate(language = factor(language, lang_order),
         model = factor(model, model_order)) %>%
  arrange(model, language) %>%
  select(model, language, auc_mean) %>%
  mutate(auc_mean = round(auc_mean, 3)) %>%
  pivot_wider(names_from = language, values_from = auc_mean) %>%
  kable(align = c("l", rep("r", length(lang_order)))) %>%
  kable_styling(full_width = FALSE)
```

## Key contrasts

Each value is LL(first model) − LL(second model) on the held-out pairs, in
nats. Positive values favour the first model. The mean and SD are over splits,
and the last column is the percentage of splits favouring the first model.
**The SD measures how much the result depends on the random split. It is not
a standard error, and these numbers are not significance tests** (see
Limitations).

```{r table-contrasts}
con_tab <- cv_con %>%
  mutate(language = factor(language, lang_order),
         contrast = factor(contrast, contrast_order)) %>%
  arrange(contrast, language) %>%
  mutate(cell = paste0(fmt(diff_mean), " ± ", fmt(diff_sd),
                       " [", round(pct_favour_first), "%]"))

write.csv(con_tab %>% select(-cell), file.path(out_dir, "table_contrasts.csv"),
          row.names = FALSE)

con_tab %>%
  select(contrast, language, cell) %>%
  pivot_wider(names_from = language, values_from = cell) %>%
  kable(align = c("l", rep("r", length(lang_order))),
        caption = "Held-out LL difference: mean ± SD over splits [% of splits favouring the first model]") %>%
  kable_styling(full_width = FALSE)
```

# 2. Leave one language out

For each language, the mixed model `SameName ~ predictors + (1 | Language)`
was fitted to the other six languages and scored on the held-out one. There
are two scoring variants:

- **Population intercept.** The model is used as is, with the random intercept
  set to zero. This is the strictest test, but the result also depends on how
  well the base rate of shared names transfers to the new language.
- **Recalibrated intercept.** The slopes come from the six training languages,
  and only the intercept is refitted on the held-out language. This asks
  whether the *relationship* between distance and name-sharing transfers. The
  same single parameter is refitted for every model.

Differences are given per 1000 held-out pairs, so that languages of different
sizes are comparable.

```{r table-lolo}
lolo_tab <- lolo_con %>%
  mutate(held_out = factor(held_out, languages),
         contrast = factor(contrast, contrast_order),
         score = recode(score, ll_pop = "Population intercept",
                        ll_recal = "Recalibrated intercept")) %>%
  arrange(score, contrast, held_out)

write.csv(lolo_tab, file.path(out_dir, "table_lolo.csv"), row.names = FALSE)

lolo_tab %>%
  select(score, contrast, held_out, diff_per_1000) %>%
  mutate(diff_per_1000 = round(diff_per_1000, 2)) %>%
  pivot_wider(names_from = held_out, values_from = diff_per_1000) %>%
  left_join(lolo_tab %>%
              distinct(score, contrast, n_favour_first, n_languages,
                       sign_test_p, mean_diff_per_1000, ci_lo, ci_hi) %>%
              mutate(`Favour first` = paste0(n_favour_first, "/", n_languages),
                     `Mean [95% CI]` = paste0(round(mean_diff_per_1000, 2),
                                              " [", round(ci_lo, 2), ", ",
                                              round(ci_hi, 2), "]"),
                     `Sign test p` = signif(sign_test_p, 2)) %>%
              select(score, contrast, `Favour first`, `Mean [95% CI]`,
                     `Sign test p`),
            by = c("score", "contrast")) %>%
  kable(caption = "LOLO: held-out LL difference per 1000 pairs (first − second model)") %>%
  kable_styling(full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```

# 3. Figure

The figure shows each contrast in each language, in nats per 1000 held-out
pairs. Filled circles are species-held-out CV, with bars for ±1 SD over
splits. Open triangles are LOLO with a recalibrated intercept. "All" is the
pooled species-held-out model.

```{r figure, fig.width = 10, fig.height = 4.5}
fig_species <- cv %>%
  cross_join(bind_rows(lapply(names(contrasts), function(cn)
    tibble(contrast = cn, a = contrasts[[cn]][1], b = contrasts[[cn]][2])))) %>%
  filter(model == a | model == b) %>%
  mutate(sign = if_else(model == a, 1, -1)) %>%
  group_by(language, contrast, rep) %>%
  summarise(diff = 1000 * sum(sign * ll) / first(n_pairs), .groups = "drop") %>%
  group_by(language, contrast) %>%
  summarise(mean = mean(diff), sd = sd(diff), .groups = "drop") %>%
  mutate(method = "Species held out")

fig_lolo <- lolo_tab %>%
  filter(score == "Recalibrated intercept") %>%
  transmute(language = as.character(held_out), contrast = as.character(contrast),
            mean = diff_per_1000, sd = NA_real_,
            method = "Language held out")

fig_levels <- rev(lang_order)
fig_data <- bind_rows(fig_species, fig_lolo) %>%
  mutate(contrast = factor(contrast, contrast_order),
         method = factor(method, c("Species held out", "Language held out")),
         # offset the LOLO marker below the species-held-out one
         y = match(language, fig_levels) +
           if_else(method == "Language held out", -0.22, 0.12))

p <- ggplot(fig_data, aes(x = mean, y = y)) +
  geom_vline(xintercept = 0, colour = "grey60", linewidth = 0.4) +
  geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd),
                 data = filter(fig_data, !is.na(sd)),
                 height = 0, colour = "grey20", linewidth = 0.5) +
  geom_point(aes(shape = method), size = 2.4, colour = "grey10",
             fill = "grey10", stroke = 0.7) +
  scale_shape_manual(values = c("Species held out" = 21,
                                "Language held out" = 2), name = NULL) +
  scale_y_continuous(breaks = seq_along(fig_levels), labels = fig_levels,
                     expand = expansion(add = 0.5)) +
  facet_wrap(~ contrast, nrow = 1, scales = "free_x") +
  labs(x = "Held-out LL difference per 1000 pairs (positive favours first model)",
       y = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top",
        panel.spacing.x = unit(1.4, "lines"),
        panel.border = element_rect(colour = "grey85", fill = NA),
        panel.grid.major.y = element_line(colour = "grey95"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = "grey92"))
p
ggsave(file.path(out_dir, "cv_contrasts.png"), p, width = 10, height = 4.5,
       dpi = 300, bg = "white")
```

# Limitations

- **Leave-one-language-out has only 7 languages, so its inferential power is
  limited.** A two-sided sign test can't reach p < .05 unless all 7 languages
  agree, and even then p ≈ .016. If 6 of 7 agree, p ≈ .125. A paired t-test
  across languages has 6 df and low power. The 7 languages are also a
  convenience sample, not a random sample of the world's folk taxonomies: they
  come from different ethnographers and are clustered by region. So "the
  relationship transfers to a new language" means "to languages like these".
  The LOLO results are therefore best read as effect sizes, language by
  language, rather than as a test.
- **The species-held-out CV isn't a significance test either.** The SD across
  splits shows how much the result depends on the random fold assignment, not
  the sampling uncertainty in the comparison. Pairs still share species and
  categories, so no simple standard error applies.
- **What the CV does provide** is a fairer point estimate of how well each
  model predicts unseen species than in-sample BIC, which treats about 70k
  dependent pairs as independent. Valid p-values for "does this measure add
  anything?" and "is 1974 better than 2023?" need the permutation tests and
  MRQAP described in `README_analysis_plan.md` §2, which are not implemented
  here.
